### 2017-01-20 CO750-ApproximationAlgorithmsTopics/lectures/lecture_04.tex
partial lecture 5 scribe

\documentclass[letterpaper,12pt,oneside,onecolumn]{article}
\usepackage[margin=1in, bottom=1in, top=1in]{geometry} %1 inch margins
\usepackage{amsmath, amssymb, amstext}
\usepackage{fancyhdr}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{theorem}
\usepackage{tikz}
\usepackage{tkz-berge}

%Macros
\newcommand{\A}{\mathbb{A}} \newcommand{\C}{\mathbb{C}}
\newcommand{\D}{\mathbb{D}} \newcommand{\F}{\mathbb{F}}
\newcommand{\N}{\mathbb{N}} \newcommand{\R}{\mathbb{R}}
\newcommand{\T}{\mathbb{T}} \newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}

\newcommand{\cA}{\mathcal{A}} \newcommand{\cB}{\mathcal{B}}
\newcommand{\cC}{\mathcal{C}} \newcommand{\cD}{\mathcal{D}}
\newcommand{\cE}{\mathcal{E}} \newcommand{\cF}{\mathcal{F}}
\newcommand{\cG}{\mathcal{G}} \newcommand{\cH}{\mathcal{H}}
\newcommand{\cI}{\mathcal{I}} \newcommand{\cJ}{\mathcal{J}}
\newcommand{\cK}{\mathcal{K}} \newcommand{\cL}{\mathcal{L}}
\newcommand{\cM}{\mathcal{M}} \newcommand{\cN}{\mathcal{N}}
\newcommand{\cO}{\mathcal{O}} \newcommand{\cP}{\mathcal{P}}
\newcommand{\cQ}{\mathcal{Q}} \newcommand{\cR}{\mathcal{R}}
\newcommand{\cS}{\mathcal{S}} \newcommand{\cT}{\mathcal{T}}
\newcommand{\cU}{\mathcal{U}} \newcommand{\cV}{\mathcal{V}}
\newcommand{\cW}{\mathcal{W}} \newcommand{\cX}{\mathcal{X}}
\newcommand{\cY}{\mathcal{Y}} \newcommand{\cZ}{\mathcal{Z}}

\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}

\newenvironment{proof}{{\bf Proof:  }}{\hfill\rule{2mm}{2mm}}
\newenvironment{proofof}[1]{{\bf Proof of #1:  }}{\hfill\rule{2mm}{2mm}}
\newenvironment{proofofnobox}[1]{{\bf#1:  }}{}\newenvironment{example}{{\bf Example:  }}{\hfill\rule{2mm}{2mm}}

%\renewcommand{\thesection}{\lecnum.\arabic{section}}
%\renewcommand{\theequation}{\thesection.\arabic{equation}}
%\renewcommand{\thefigure}{\thesection.\arabic{figure}}

\newtheorem{fact}{Fact}[section]
\newtheorem{lemma}[fact]{Lemma}
\newtheorem{theorem}[fact]{Theorem}
\newtheorem{definition}[fact]{Definition}
\newtheorem{corollary}[fact]{Corollary}
\newtheorem{proposition}[fact]{Proposition}
\newtheorem{claim}[fact]{Claim}
\newtheorem{exercise}[fact]{Exercise}
\newtheorem{note}[fact]{Note}
\newtheorem{conjecture}[fact]{Conjecture}

\newcommand{\size}[1]{\ensuremath{\left|#1\right|}}
\newcommand{\ceil}[1]{\ensuremath{\left\lceil#1\right\rceil}}
\newcommand{\floor}[1]{\ensuremath{\left\lfloor#1\right\rfloor}}

%END MACROS
%Page style
\pagestyle{fancy}

\listfiles

\raggedbottom

\lhead{2017-01-17}
\rhead{William Justin Toth CO750-Approximation Algorithms Lecture 4} %CHANGE n to ASSIGNMENT NUMBER ijk TO COURSE CODE
\renewcommand{\headrulewidth}{1pt} %heading underlined
%\renewcommand{\baselinestretch}{1.2} % 1.2 line spacing for legibility (optional)

\begin{document}
Recall from Lecture $3$:
\begin{theorem}
Given any metric $(V,d)$ one can randomly find a tree $T$ such that
\begin{enumerate}
\item $d(u,v) \leq d^T(u,v)$ for all $u,v\in V$
\item $E[d^T(u,v)] \leq O(\log n) d(u,v)$ for all $u,v\in V$.
\end{enumerate}
\end{theorem}
\begin{proof}
We may assume $\max_{u,v \in V} d(u,v) = 2^\delta$ for some $\delta \in \N$ and $d(u,v) > 1$ for all $u,v\in V$. Construct a tree $T$ by:
\begin{enumerate}
\item Choose randomly a permutation $\pi$ of $V$.
\item Choose $\beta \in [0,1]$ uniformly at random.
\item Set $D_\delta = \{V\}$.
\item For $i = \delta -1$ to $0$ do:
\begin{enumerate}
\item Assign every node to the first node according to $\pi$ that has distance at most $2^\beta \cdot 2^{i-1}$.
\item All nodes assigned to the same node that are currently in the same cluster will form a new cluster in $D_i$.
\end{enumerate}
\item Return tree $T$ where $V(T) = \{s \in D_i : \forall i = 0, \dots, \delta\}$ and $E(T) = \{s_is_{i-1}  \in D_{i}\times D_{i-1}: s_{i-1} \subseteq s_i, \forall i = 0,\dots,\delta\}$, and $d(s_is_{i-1}) = 2^i$.
\end{enumerate}
Note that the clusters of $D_0$ are singletons because the bound on the maximum distance is $2^\beta 2^0 \leq 1$. Hence the clusters of $D_0$ can be identified with nodes in $V$. 
\paragraph{Proof of 1.}
Suppose $u,v$ are in the same cluster at level $D_i$ and are separated at level $D_{i-1}$. Let $w$ be the node identified with $u$ and $v$'s cluster $S_w$ at level $D_i$. Then $d(u,v) \leq d(u,w) + d(w,v) \leq 2^\beta 2^{i-1} + 2^\beta 2^{i-1} \leq 2^{i+1}$. On the other hand $d^T(u, S_w) \geq 2^i$ and $d^T(S_w, v) \geq 2^i$ so $d^T(u,v) \geq 2^i + 2^i = 2^{i+1}$. This proves part $1.$ of the theorem.
\paragraph{Proof of 2.}
Observe that if $u,v$ are separated first at level $D_i$ then $d^T(u,v) \leq 2^{i+2}$. Let $S_i$ be the cluster in $D_i$ containing $u,v$. Let $S^u_{i-1}$ be descended cluster of $S_i$ from which $u$ descends and similarly define $S^v_{i-1}$. Then 
\begin{align*}
d^T(u,v) &= d^T(S_i, S^u_{i-1}) + d^T(S_, S^u_{i-1}) + d^T(S^u_{i-1}, u) + d^T(S^v_{i-1}, v)\\
&= 2^i + 2^i +  d^T(S^u_{i-1}, u) + d^T(S^v_{i-1}, v) \\
&\leq 2^i + 2^i + 2\sum_{j = 1}^{i-1} 2^j \\
&= 2^{i+1} + 2^{i+1} \\
&= 2^{i+2}.
\end{align*}
Now we way that node $c$ {\it settles} $(u,v)$ at level $D_i$ if $c$ is the first node with respect to $\pi$ to which one of $u$ or $v$ gets assigned at level $D_i$. We say that node $c$ {\it cuts} $(u,v)$ at level $D_i$ if exactly one of $u,v$ is at distance at most $2^\beta 2^i$ from $c$. If $u,v$ are separated at level $D_i$ then there exists some $c$ which both settles and cuts $u,v$ at level $D_i$.
\paragraph{}
Note that $$E[d^T(u,v)] \leq \sum_{i=0}^{\delta - 1} \sum_{c \in V} Pr [ \text{$c$ settles (A) and cuts (B) $u,v$ at $D_i$}] 2^{i+2}.$$ The bound simply follows from our observation, and the prior definitions. Recall that $Pr[A \cap B] = P[A\mid B] Pr[B]$.
\paragraph{}
We first analyze $Pr[B]$. Without loss we may assume $d(c,u) \leq d(c,v)$. For $c$ to cut $u,v$ we need $d(c,u) \leq 2^\beta 2^{i-1} < d(c,v)$ by definition. Thus
$$Pr[B] = \frac{| [2^{i-1},2^i] \cap [d(c,u),d(c,v)]|}{|[2^{i-1},2^i]|} = \frac{| [2^{i-1},2^i] \cap [d(c,u),d(c,v)]|}{2^{i-1}}$$.
Now we need $Pr[A \mid B]$. Let us order the nodes of $V$ according to their distance to $u,v$. That is, $V = \{c_1, \dots, c_n\}$ where $\min\{d(c_i, v), d(c_i, u)\} \leq \min \{d(c_{i+1}, v), d(c_{i+1}, u)\}$ for all $i$. If $c= c_k$ then the probability that $c$ comes before $c_1, \dots, c_{k-1}$ in $\pi$ is at most $\frac{1}{k}$. In order for $c$ (given that $c$ cuts $u,v$) to settle $u,v$ at least that has to happen. Hence
$$Pr[A \mid B] \leq \frac{1}{k}.$$
\paragraph{}
Putting it all together we obtain the following expectation:
\begin{align*}
E[d^T(u,v)] &\leq \sum_{i=0}^{\delta -1} \sum_{k=1}^n Pr [A \mid B]Pr[B] 2^{i+2} \\
&\leq \sum_{k=1}^n \frac{1}{k} \sum_{i=0}^{\delta -1 }  2^{i+2}\frac{| [2^{i-1},2^i] \cap [d(c,u),d(c,v)]|}{2^{i-1}} \\
&= 8 \sum_{k=1}^n \frac{1}{k} \sum_{i=0}^{\delta -1 }  | [2^{i-1},2^i] \cap [d(c,u),d(c,v)]| \\
&\leq 8 d(u,v) \sum_{k=1}^n \frac{1}{k} \\
&= 8 H_n d(u,v) \\
&= O(log n) d(u,v).
\end{align*}
Thus the second claim and therefore the theorem holds.
\end{proof}
\begin{theorem}
There exists a tree $T' = (V,F)$ such that $d^T(u,v) \leq d^{T'}(u,v) \leq 4 d^T(u,v).$
\end{theorem}
\begin{proof}
Pick $v \in V$ with a parent $w \not \in V$ and contract edge $vw$ in $T$. Repeat until all nodes of $T$ are in $V$. Multiply by $4$ the weights on all remaining edges. Call the resulting tree $T'$. Note that clearly $d^{T'}(u,v) \leq 4d^T(u,v)$ as contracting only shortens the distances. Furthermore, in the worst case $d^{T'}(u,v) = 2^i \cdot 4$ if contraction results in contracting $u$ and $v$ right up to the cluster where they were separated (that is, contracting preserves the largest edge distance on the $uv$-path in $T$). But 
$$2^i \cdot 4 = 2^i + 2^i + 2^i + 2^i \geq d(S^u_{i-1}, S_i) + d(S^v_{i-1}, S_i) +  d(S^u_{i-1}, u) + d(S^v_{i-1}, v) = d^T(u,v).$$
Hence the theorem holds.
\end{proof}
\section{Generalized Steiner Network}
\paragraph{}
Here we seek to given an application of the previous theory. In the Generalized Steiner Network problem (GSN), we are given a graph $G=(V,E)$, a distance function $d: E \rightarrow \R_+$, and requirements $r(u,v) \in \N$ for all $u,v \in V$. We are to find an integer capacity function $x : E \rightarrow \N$ of minimum cost $\sum_{e \in E} d(e) x_e$ such that capacities $x$ allow for routing a flow of value $r(u,v)$ from $u$ to $v$.
\paragraph{Proposition 1} Optimizing on $G$ is equivalent to optimizing on $G^M$ for $GSN$ problems. This proposition is immediate from the same arguments as was used for ST.
\paragraph{Proposition 2} GSN is solvable in polynomial time on trees. Since each edge is a cut in the tree, we must set $x_{uv} = r(u,v)$ and this is the best we can do.
\paragraph{Algorithm} Consider the following simple algorithm for GSN:
\begin{enumerate}
\item Find a tree $T$ of $G$ according to previous section's metric embedding algorithms.
\item Solve GSN on $T$ and return the solution.
\end{enumerate}
\begin{theorem}
Algorithm above is a randomized $O(log n)$-approximation algorithm for GSN.
\end{theorem}
\begin{proof}
Let $\bar{x}$ be the solution return by our algorithm. Let $x^*$ be OPT solution. Then $\sum_{uv} d(u,v) \bar{x}_{uv} \leq \sum_{uv} d^T(u,v)\bar{x}_{uv} \leq \sum_{uv} d^T(u,v) x^*_{uv}$. The first inequality follows since $d(u,v) \leq d^T(u,v)$. The second inequality follows since $x^*$ is a feasible solution for $T$. Taking expectation we see:
$$E[\sum_{uv} d(u,v) \bar{x}_{uv}] \leq E[\sum_{uv} d^T(u,v)] \leq O(log n) \sum_{uv} d(u,v) x^*_{uv}$$
with the last inequality following from our theorem.
\end{proof}
\end{document}