### 2017-01-28 CO750-ApproximationAlgorithmsTopics/assignments/assignment_01.tex
1a apxalgs, 1,2a convopt

\documentclass[letterpaper,12pt,oneside,onecolumn]{article}
\usepackage[margin=1in, bottom=1in, top=1in]{geometry} %1 inch margins
\usepackage{amsmath, amssymb, amstext}
\usepackage{fancyhdr}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{theorem}
\usepackage{tikz}
\usepackage{tkz-berge}

%Macros
\newcommand{\A}{\mathbb{A}} \newcommand{\C}{\mathbb{C}}
\newcommand{\D}{\mathbb{D}} \newcommand{\F}{\mathbb{F}}
\newcommand{\N}{\mathbb{N}} \newcommand{\R}{\mathbb{R}}
\newcommand{\T}{\mathbb{T}} \newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}

\newcommand{\cA}{\mathcal{A}} \newcommand{\cB}{\mathcal{B}}
\newcommand{\cC}{\mathcal{C}} \newcommand{\cD}{\mathcal{D}}
\newcommand{\cE}{\mathcal{E}} \newcommand{\cF}{\mathcal{F}}
\newcommand{\cG}{\mathcal{G}} \newcommand{\cH}{\mathcal{H}}
\newcommand{\cI}{\mathcal{I}} \newcommand{\cJ}{\mathcal{J}}
\newcommand{\cK}{\mathcal{K}} \newcommand{\cL}{\mathcal{L}}
\newcommand{\cM}{\mathcal{M}} \newcommand{\cN}{\mathcal{N}}
\newcommand{\cO}{\mathcal{O}} \newcommand{\cP}{\mathcal{P}}
\newcommand{\cQ}{\mathcal{Q}} \newcommand{\cR}{\mathcal{R}}
\newcommand{\cS}{\mathcal{S}} \newcommand{\cT}{\mathcal{T}}
\newcommand{\cU}{\mathcal{U}} \newcommand{\cV}{\mathcal{V}}
\newcommand{\cW}{\mathcal{W}} \newcommand{\cX}{\mathcal{X}}
\newcommand{\cY}{\mathcal{Y}} \newcommand{\cZ}{\mathcal{Z}}

\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}

\newenvironment{proof}{{\bf Proof:  }}{\hfill\rule{2mm}{2mm}}
\newenvironment{proofof}[1]{{\bf Proof of #1:  }}{\hfill\rule{2mm}{2mm}}
\newenvironment{proofofnobox}[1]{{\bf#1:  }}{}\newenvironment{example}{{\bf Example:  }}{\hfill\rule{2mm}{2mm}}

%\renewcommand{\thesection}{\lecnum.\arabic{section}}
%\renewcommand{\theequation}{\thesection.\arabic{equation}}
%\renewcommand{\thefigure}{\thesection.\arabic{figure}}

\newtheorem{fact}{Fact}[section]
\newtheorem{lemma}[fact]{Lemma}
\newtheorem{theorem}[fact]{Theorem}
\newtheorem{definition}[fact]{Definition}
\newtheorem{corollary}[fact]{Corollary}
\newtheorem{proposition}[fact]{Proposition}
\newtheorem{claim}[fact]{Claim}
\newtheorem{exercise}[fact]{Exercise}
\newtheorem{note}[fact]{Note}
\newtheorem{conjecture}[fact]{Conjecture}

\newcommand{\size}[1]{\ensuremath{\left|#1\right|}}
\newcommand{\ceil}[1]{\ensuremath{\left\lceil#1\right\rceil}}
\newcommand{\floor}[1]{\ensuremath{\left\lfloor#1\right\rfloor}}

%END MACROS
%Page style
\pagestyle{fancy}

\listfiles

\raggedbottom

\lhead{2017-02-09}
\rhead{William Justin Toth CO750-Approximation Algorithms Assignment 1} %CHANGE n to ASSIGNMENT NUMBER ijk TO COURSE CODE
\renewcommand{\headrulewidth}{1pt} %heading underlined
%\renewcommand{\baselinestretch}{1.2} % 1.2 line spacing for legibility (optional)

\begin{document}
\section{}
\paragraph{}
Let $E$ be the edge set of a graph $G$. Let $\cS$ be the collection of all sets $S \subseteq E$ that are matchings of $G$. Let $w: E \rightarrow \R_+$ be a weight function on the edges. For a subset $\bar{E} \subseteq E$ we denote by $G[\bar{E}]$ the subgraph of $G$ induced by the edges in $\bar{E}$.
\subsection{a}
\paragraph{}
Let $\bar{E} \subseteq E$. Suppose that $M \subseteq \bar{E}$ is an inclusion-wise maximal matching of $G[\bar{E}]$. Let $\bar{M}$ be a matching of $G[\bar{E}]$. Let $f : M \rightarrow \{ \{e_1,e_2\} | e_1, e_2 \in \bar{M}\}$ be defined by
$$ f(uv) = \{e_1, e_2\}$$
where $e_1 \in \bar{M}$ covers $u$ and $e_2 \in \bar{M}$ covers $v$, for each $uv \in M$.
\paragraph{}
We claim that for all $uv \in \bar{M}$, there exists $P \in Im(f)$ (where $Im(f)$ denotes the image of $f$) such that $uv \in P$. To see this suppose for a contradiction that there exists some $uv \in \bar{M}$ where each $P \in Im(f)$ does not contain $uv$. Then by the definition of $f$, neither $u$ nor $v$ are covered by $M$. But this violates that $M$ is inclusion-wise maximal since $uv \not\in M$ and $M \cup \{uv\}$ is a matching. Hence the claim holds.
\paragraph{}
From the previous claim we observe $|\bar{M}| \leq | \cup_{P \in Im(f)} P|$. But this yields a string of inequalities which complete the proof:
\begin{align*}
|\bar{M}| &\leq |\cup_{P \in Im(f)} P| \\
&\leq \sum_{P \in Im(f)} |P| \\
&\leq \sum_{P \in Im(f)} 2 \\
&\leq 2|Im(f)| \\
&\leq 2|M|. 
\end{align*}
With the last inequality following since $M$ is the domain of $f$. Hence we have $|M| \geq \frac{1}{2}|\bar{M}|$ as desired. $\blacksquare$
\subsection{b}
\paragraph{}
Let $e$
\end{document}