### 2017-01-30 CO750-ApproximationAlgorithmsTopics/assignments/assignment_01.tex
Fixed 3b of apx algs a1

\documentclass[letterpaper,12pt,oneside,onecolumn]{article}
\usepackage[margin=1in, bottom=1in, top=1in]{geometry} %1 inch margins
\usepackage{amsmath, amssymb, amstext}
\usepackage{fancyhdr}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{theorem}
\usepackage{tikz}
\usepackage{tkz-berge}

%Macros
\newcommand{\A}{\mathbb{A}} \newcommand{\C}{\mathbb{C}}
\newcommand{\D}{\mathbb{D}} \newcommand{\F}{\mathbb{F}}
\newcommand{\N}{\mathbb{N}} \newcommand{\R}{\mathbb{R}}
\newcommand{\T}{\mathbb{T}} \newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}

\newcommand{\cA}{\mathcal{A}} \newcommand{\cB}{\mathcal{B}}
\newcommand{\cC}{\mathcal{C}} \newcommand{\cD}{\mathcal{D}}
\newcommand{\cE}{\mathcal{E}} \newcommand{\cF}{\mathcal{F}}
\newcommand{\cG}{\mathcal{G}} \newcommand{\cH}{\mathcal{H}}
\newcommand{\cI}{\mathcal{I}} \newcommand{\cJ}{\mathcal{J}}
\newcommand{\cK}{\mathcal{K}} \newcommand{\cL}{\mathcal{L}}
\newcommand{\cM}{\mathcal{M}} \newcommand{\cN}{\mathcal{N}}
\newcommand{\cO}{\mathcal{O}} \newcommand{\cP}{\mathcal{P}}
\newcommand{\cQ}{\mathcal{Q}} \newcommand{\cR}{\mathcal{R}}
\newcommand{\cS}{\mathcal{S}} \newcommand{\cT}{\mathcal{T}}
\newcommand{\cU}{\mathcal{U}} \newcommand{\cV}{\mathcal{V}}
\newcommand{\cW}{\mathcal{W}} \newcommand{\cX}{\mathcal{X}}
\newcommand{\cY}{\mathcal{Y}} \newcommand{\cZ}{\mathcal{Z}}

\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}

\newenvironment{proof}{{\bf Proof:  }}{\hfill\rule{2mm}{2mm}}
\newenvironment{proofof}[1]{{\bf Proof of #1:  }}{\hfill\rule{2mm}{2mm}}
\newenvironment{proofofnobox}[1]{{\bf#1:  }}{}\newenvironment{example}{{\bf Example:  }}{\hfill\rule{2mm}{2mm}}

%\renewcommand{\thesection}{\lecnum.\arabic{section}}
%\renewcommand{\theequation}{\thesection.\arabic{equation}}
%\renewcommand{\thefigure}{\thesection.\arabic{figure}}

\newtheorem{fact}{Fact}[section]
\newtheorem{lemma}[fact]{Lemma}
\newtheorem{theorem}[fact]{Theorem}
\newtheorem{definition}[fact]{Definition}
\newtheorem{corollary}[fact]{Corollary}
\newtheorem{proposition}[fact]{Proposition}
\newtheorem{claim}[fact]{Claim}
\newtheorem{exercise}[fact]{Exercise}
\newtheorem{note}[fact]{Note}
\newtheorem{conjecture}[fact]{Conjecture}

\newcommand{\size}[1]{\ensuremath{\left|#1\right|}}
\newcommand{\ceil}[1]{\ensuremath{\left\lceil#1\right\rceil}}
\newcommand{\floor}[1]{\ensuremath{\left\lfloor#1\right\rfloor}}

%END MACROS
%Page style
\pagestyle{fancy}

\listfiles

\raggedbottom

\lhead{2017-02-09}
\rhead{William Justin Toth CO750-Approximation Algorithms Assignment 1} %CHANGE n to ASSIGNMENT NUMBER ijk TO COURSE CODE
\renewcommand{\headrulewidth}{1pt} %heading underlined
%\renewcommand{\baselinestretch}{1.2} % 1.2 line spacing for legibility (optional)

\begin{document}
%Q1
\section{}
\paragraph{}
Let $E$ be the edge set of a graph $G$. Let $\cS$ be the collection of all sets $S \subseteq E$ that are matchings of $G$. Let $w: E \rightarrow \R_+$ be a weight function on the edges. For a subset $\bar{E} \subseteq E$ we denote by $G[\bar{E}]$ the subgraph of $G$ induced by the edges in $\bar{E}$.
	%Pa
\subsection{a}
\paragraph{}
Let $\bar{E} \subseteq E$. Suppose that $M \subseteq \bar{E}$ is an inclusion-wise maximal matching of $G[\bar{E}]$. Let $\bar{M}$ be a matching of $G[\bar{E}]$. Let $f : M \rightarrow \{ \{e_1,e_2\} | e_1, e_2 \in \bar{M}\}$ be defined by
$$ f(uv) = \{e_1, e_2\}$$
where $e_1 \in \bar{M}$ covers $u$ and $e_2 \in \bar{M}$ covers $v$, for each $uv \in M$.
\paragraph{}
We claim that for all $uv \in \bar{M}$, there exists $P \in Im(f)$ (where $Im(f)$ denotes the image of $f$) such that $uv \in P$. To see this suppose for a contradiction that there exists some $uv \in \bar{M}$ where each $P \in Im(f)$ does not contain $uv$. Then by the definition of $f$, neither $u$ nor $v$ are covered by $M$. But this violates that $M$ is inclusion-wise maximal since $uv \not\in M$ and $M \cup \{uv\}$ is a matching. Hence the claim holds.
\paragraph{}
From the previous claim we observe $|\bar{M}| \leq | \cup_{P \in Im(f)} P|$. But this yields a string of inequalities which complete the proof:
\begin{align*}
|\bar{M}| &\leq |\cup_{P \in Im(f)} P| \\
&\leq \sum_{P \in Im(f)} |P| \\
&\leq \sum_{P \in Im(f)} 2 \\
&\leq 2|Im(f)| \\
&\leq 2|M|. 
\end{align*}
With the last inequality following since $M$ is the domain of $f$. Hence we have $|M| \geq \frac{1}{2}|\bar{M}|$ as desired. $\blacksquare$
	%Pb
\subsection{b}
\paragraph{}
Let $e$

%Q2
\section{}

%Q3
\section{}
\paragraph{}
Let $G = (V,E)$ be a graph. For any $S, T \subseteq V$ we denote by $E(S,T)$ the set of edges between $S$ and $T$. Formally $E(S,T) = \{\{s,t\} \in E: s \in S, t \in T\}$.
%Pa
\subsection{a}

%Pb
\subsection{b}
\paragraph{}
Our goal is to find a maximum $k$-Cut of $G$. That is, we seek to find a partition of $V$ into $V_1, \dots, V_k$ maximizing
$$\sum_{i=1}^k \sum_{j = 1}^{i-1} |E(V_i, V_j)| := c(\{V_1, \dots, V_k\}).$$
Intuitively we are maximizing the edges between the selected cuts. Let $\cB$ denote the $2$-approximation algorithm for max cut given in problem $3a$. Consider the following algorithm, which we will denote by $\cA$:
\begin{enumerate}
\item Set $V_1 = S$ and $V_2 = V\backslash S$ where $S$ is the cut returned by $\cB$ when run on $G$.
\item Set $V_i = \emptyset$ for $i = 3, \dots, k$.
\item For $i = 2, \dots, k-1$
\begin{enumerate}
\item For $j = 1, \dots, i$ set $S_j$ to the cut returned by $\cB$ when run on $G[V_j]$.
\item Set $j^* = \text{argmax}_{j=1,\dots, i} \{|E(S_j, V_j\backslash S_j)|\}$.
\item Set $V_{i+1} = S_{j^*}$ and set $V_{j^*} = V_{j^*} \backslash S_{j^*}$.
\end{enumerate}
\item Return $V_1, \dots, V_k$.
\end{enumerate}
The idea behind the operation of $\cA$ is to start with a $2$-cut of $G$, and in every successive iteration take the current partition and $2$-cut one of the parts in an approximately optimal way until we reach a $k$-cut.
\begin{lemma}\label{lemma:3b1}
Let $\cO = \{V^*_1, \dots, V^*_k\}$. Consider an arbitrary iteration $i = 1, \dots, k-1$ (iteration $1$ is original $2$-cut). Let $\cS = \{V_1, \dots, V_{i+1}\}$ be the set of partitions chosen so far by $\cA$. Suppose that at the end of iteration $i$, the algorithm has split $V_j$ into $V_j'$ and $V_{i+1}$. That is the solution after iteration $i$ is
$$\cS' = \cS \backslash V_j \cup \{V_j', V_{i+1}\}.$$
Then 
$$2k(c(\cS') - c(\cS)) \geq c(\cO) - c(\cS).$$
\end{lemma}
\begin{proof}
The difference between $c(\cS')$ and $c(\cS)$ is precisely $|E(V_j', V_{i+1})|$ (the number of edges gained by splitting $V_j$). Further the difference $c(\cO) - c(\cS)$ is upper-bounded by the edges counted in $c(\cO)$ that are not counted in $c(\cS)$. Call such edges $E'$.  The edges $E'$ are necessarily internal to a part of $\cS$ (that is, both endpoints lie in some $V_k \in \cS$). Otherwise they are counted in $c(\cS)$ since their endpoints lie in distinct parts of $\cS$. Thus if we consider the $k$-cut induced by $\cO$ on $G[E']$ it consists of at most $k$ splits ($2$-cuts) of parts $V_k \in \cS$.
\paragraph{}
Let $V'_k$ denote the best possible $2$-cut of $V_k \in \cS$ for all $k = 1, \dots, i$. Now let $$k^* = \text{argmax}_{k=1,\dots,i}  \{|E(V'_k, V_k \backslash V'_k)|\}.$$ We have
$$ |E'| \leq k |E(V'_{k^*}, V_k \backslash V'_{k^*})|$$
by our previous discussion, since $E'$ edges are formed by a $k$-cut on edges internal to parts of $\cS$, and  $|E(V'_{k^*}, V_k \backslash V'_{k^*}|$ is largest $2$-cut to be formed by edges internal to a part of $\cS$. Now the cut found by our algorithm $\cA$ in this iteration in step $3(c)$ is a $2$-approximation of the best possible split cut. That is,
$$|E(V'_{k^*}, V_k \backslash V'_{k^*})| \leq 2|E(V_j', V_{i+1})| .$$
Hence, combining these inequalities, we see that
$$|E'| \leq 2k |E(V_j', V_{i+1})|.$$
Therefore $2k(c(\cS') - c(\cS)) \geq c(\cO) - c(\cS)$ as desired.
\end{proof}
\paragraph{}
Using the previous lemma, we can demonstrate that $\cA$ is a $\alpha := 1-(1-\frac{1}{2k})^k$-approximation algorithm for the max $k$-cut problem.
\paragraph{Main Proof}
Since $\cB$ can be run in polynomial time, and $k$ is at most $|V|$, it is clear that $\cA$ runs in polynomial time. At termination the algorithm returns a partition of $V$ into $k$ parts so it returns a feasible solution.
\paragraph{} 
Now to see the approximation factor holds let $\cS^i$ denote the solution maintained by $\cA$ at the end of iteration $i$. Then the returned solution is $\cS^{k-1}$. Let $\cO$ be an optimal solution. Observe from Lemma \ref{lemma:3b1} that $c(\cS^{i}) \geq \frac{1}{2k} c(\cO) + (1- \frac{1}{2k})c(S^{i-1})$. We will apply this inequality inductively to achieve the bound:
\begin{align*}
c(\cS^{k-1} &\geq \frac{1}{2k} c(\cO) + (1- \frac{1}{2k})c(S^{k-2}) \\
		&\geq \frac{1}{2k} c(\cO) + (1- \frac{1}{2k})(\frac{1}{2k} c(\cO) + (1- \frac{1}{2k})c(S^{k-3})) \\
		&\geq \frac{1}{2k} c(\cO) + (1- \frac{1}{2k})(1 + (1-\frac{1}{2k}) + (1-\frac{1}{2k})^2 + \dots + (1-\frac{1}{2k})^{k-1}) \\
		&= \frac{c(\cO)}{2k} \cdot \frac{1-(1-\frac{1}{2k})^k}{1-(1-\frac{1}{2k})} \\
		&=\frac{c(\cO)}{2k} \cdot \frac{1-(1-\frac{1}{2k})^k}{\frac{1}{2k}}\\
		&= 1-(1-\frac{1}{2k})^kc(\cO).
\end{align*}
Thus the approximation factor holds as desired. $\blacksquare$
\end{document}