### 2017-03-28 CS798-ConvexityAndOptimization/assignments/assignment_02.tex
writeup q3

\documentclass[letterpaper,12pt,oneside,onecolumn]{article}
\usepackage[margin=1in, bottom=1in, top=1in]{geometry} %1 inch margins
\usepackage{amsmath, amssymb, amstext}
\usepackage{fancyhdr}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{theorem}
\usepackage{tikz}
\usepackage{tkz-berge}

%Macros
\newcommand{\A}{\mathbb{A}} \newcommand{\C}{\mathbb{C}}
\newcommand{\D}{\mathbb{D}} \newcommand{\F}{\mathbb{F}}
\newcommand{\N}{\mathbb{N}} \newcommand{\R}{\mathbb{R}}
\newcommand{\T}{\mathbb{T}} \newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}

\newcommand{\cA}{\mathcal{A}} \newcommand{\cB}{\mathcal{B}}
\newcommand{\cC}{\mathcal{C}} \newcommand{\cD}{\mathcal{D}}
\newcommand{\cE}{\mathcal{E}} \newcommand{\cF}{\mathcal{F}}
\newcommand{\cG}{\mathcal{G}} \newcommand{\cH}{\mathcal{H}}
\newcommand{\cI}{\mathcal{I}} \newcommand{\cJ}{\mathcal{J}}
\newcommand{\cK}{\mathcal{K}} \newcommand{\cL}{\mathcal{L}}
\newcommand{\cM}{\mathcal{M}} \newcommand{\cN}{\mathcal{N}}
\newcommand{\cO}{\mathcal{O}} \newcommand{\cP}{\mathcal{P}}
\newcommand{\cQ}{\mathcal{Q}} \newcommand{\cR}{\mathcal{R}}
\newcommand{\cS}{\mathcal{S}} \newcommand{\cT}{\mathcal{T}}
\newcommand{\cU}{\mathcal{U}} \newcommand{\cV}{\mathcal{V}}
\newcommand{\cW}{\mathcal{W}} \newcommand{\cX}{\mathcal{X}}
\newcommand{\cY}{\mathcal{Y}} \newcommand{\cZ}{\mathcal{Z}}

\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}

\newenvironment{proof}{{\bf Proof:  }}{\hfill\rule{2mm}{2mm}}
\newenvironment{proofof}[1]{{\bf Proof of #1:  }}{\hfill\rule{2mm}{2mm}}
\newenvironment{proofofnobox}[1]{{\bf#1:  }}{}\newenvironment{example}{{\bf Example:  }}{\hfill\rule{2mm}{2mm}}

%\renewcommand{\thesection}{\lecnum.\arabic{section}}
%\renewcommand{\theequation}{\thesection.\arabic{equation}}
%\renewcommand{\thefigure}{\thesection.\arabic{figure}}

\newtheorem{fact}{Fact}[section]
\newtheorem{lemma}[fact]{Lemma}
\newtheorem{theorem}[fact]{Theorem}
\newtheorem{definition}[fact]{Definition}
\newtheorem{corollary}[fact]{Corollary}
\newtheorem{proposition}[fact]{Proposition}
\newtheorem{claim}[fact]{Claim}
\newtheorem{exercise}[fact]{Exercise}
\newtheorem{note}[fact]{Note}
\newtheorem{conjecture}[fact]{Conjecture}

\newcommand{\size}[1]{\ensuremath{\left|#1\right|}}
\newcommand{\ceil}[1]{\ensuremath{\left\lceil#1\right\rceil}}
\newcommand{\floor}[1]{\ensuremath{\left\lfloor#1\right\rfloor}}

%END MACROS
%Page style
\pagestyle{fancy}

\listfiles

\raggedbottom

\lhead{2017-03-27}
\rhead{William Justin Toth CS798-Convexity and Optimization Assignment 2} %CHANGE n to ASSIGNMENT NUMBER ijk TO COURSE CODE
\renewcommand{\headrulewidth}{1pt} %heading underlined
%\renewcommand{\baselinestretch}{1.2} % 1.2 line spacing for legibility (optional)

\begin{document}

%Question 1
\section{}

%Question 2
\section{}

%Question 3
\section{}
\paragraph{}
I collaborated with Sharat on this problem.
\paragraph{}
Let $\cA$ denote our black-box algorithm. Let $f$ be at $\beta$-smooth function. Given an initial point $x_0$ and desired accuracy $\epsilon > 0$ our algorithm does:
\begin{enumerate}
\item If $||x_0 - x^*||_2^2 = 0$ return $x_0$.
\item Let $\epsilon' = \epsilon/2$ and set $\alpha = \frac{\epsilon}{||x_0 - x^*||_2^2}$.
\item Let $g$ be the function given by $g(x) := f(x) + \frac{\alpha}{2}||x - x_0||_2^2$.
\item Run $\cA$ on $g$ with accuracy $\epsilon'$ and starting point $x_0$ returning a point $x'$.
\item Return solution $x'$.
\end{enumerate} 
Our algorithm assumes that the value of $||x_0 - x^*||_2$ is known in advance. In the trivial case where we know $x^*$ the algorithm returns it in step $1$. So we may assume that $x_0 \neq x^*$.
\begin{lemma}\label{lemma:convex-smooth}
The function $g$ is $\alpha$-strongly convex and $(\beta + \alpha)$-smooth.
\end{lemma}
\begin{proof}
The Hessian of $g$ is:
$$\nabla^2 g = \nabla^2 f + \alpha I.$$
Since we know $f$ is $\beta$-smooth we have
$$0I \preccurlyeq \nabla^2 f \preccurlyeq \beta I.$$
Since adding scalar multiples of $I$ only shifts eigenvalues we observe
$$\alpha I \preccurlyeq \nabla^2 g \preccurlyeq (\beta + \alpha) I.$$
Hence the lemma holds as desired. 
\end{proof}
\paragraph{}
By lemma \ref{lemma:convex-smooth} we know that $\cA$ can process our function $g$. It remains to verify the accuracy of the point returned and to bound the running time.
\begin{lemma}\label{lemma:q3-accuracy}
The point $x'$ returned by our algorithm satisfies
$$f(x') - f(x^*) \leq \epsilon.$$
\end{lemma}
\begin{proof}
Let $\tilde{x}$ be the optimal solution for $g$, By our guarantees on $\cA$:
$$g(x') - g(\tilde{x}) \leq \epsilon'.$$
Now observe that
\begin{align*}
f(x') - f(x^*) &= g(x') - \frac{\alpha}{2}||x'-x_0||^2_2 - g(x^*) +\frac{\alpha}{2}||x^*-x_0||_2^2 \\
&\leq g(x') - g(x^*) + \frac{\alpha}{2}||x^*-x_0||_2^2 &\text{since $\frac{\alpha}{2}||x'-x_0||^2_2\geq 0$} \\
&= g(x') - g(x^*) + \frac{\epsilon}{2} &\text{by our choice of $\alpha$} \\
&\leq g(\tilde{x}) + \epsilon' -g(x^*) + \frac{\epsilon}{2} \\
&\leq \epsilon' +\frac{\epsilon}{2} &\text{by optimality of $\tilde{x}$} \\
&= \epsilon &\text{by our choice of $\epsilon'$}.
\end{align*}
\end{proof}
\paragraph{}
Therefore our algorithm operates properly and returns a point within the desired accuracy. It remains to verify the running time of our algorithm. This running time is entirely bounded by the operation of $\cA$ on $g$ with $\epsilon'$. By the guarantee on $\cA$ this gives a bound of:
\begin{align*}
O(\sqrt{\frac{\beta}{\alpha} + 1}\log (\frac{g(x_0)-g(\tilde{x})}{\epsilon'})) &= O(\sqrt{\frac{\beta}{\alpha}}\log (\frac{g(x_0)-g(\tilde{x})}{\epsilon})) &\text{dropping constants}\\
&= O(\sqrt{\frac{\beta}{\alpha}}\log (\frac{f(x_0)-f(\tilde{x}) - \frac{\alpha}{2}||x_0-x^*||^2_2}{\epsilon})) &\text{expanding $g$}\\
&= O(\sqrt{\frac{\beta}{\alpha}}\log (\frac{f(x_0)-f(\tilde{x})}{\epsilon})) &\text{dropping non-positive term} \\
&= O(\sqrt{\frac{\beta}{\alpha}}\log (\frac{f(x_0)-f(x^*)}{\epsilon})) &\text{by optimality of $x^*$} \\
&= O(\sqrt{\frac{\beta}{\epsilon}} ||x_0-x^*||_2\log (\frac{f(x_0)-f(x^*)}{\epsilon}))  &\text{expanding $\alpha$}.
\end{align*}
Thus we obtain the desired running time. Together with lemma \ref{lemma:q3-accuracy} this completes the problem.
\end{document}