### 2017-04-06 CO750-ApproximationAlgorithmsTopics/assignments/report.tex
finished report

\documentclass[letterpaper,12pt,oneside,onecolumn]{article}
\usepackage[margin=1in, bottom=1in, top=1in]{geometry} %1 inch margins
\usepackage{amsmath, amssymb, amstext}
\usepackage{fancyhdr}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{theorem}
\usepackage{tikz}
\usepackage{tkz-berge}

%Macros
\newcommand{\A}{\mathbb{A}} \newcommand{\C}{\mathbb{C}}
\newcommand{\D}{\mathbb{D}} \newcommand{\F}{\mathbb{F}}
\newcommand{\N}{\mathbb{N}} \newcommand{\R}{\mathbb{R}}
\newcommand{\T}{\mathbb{T}} \newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}

\newcommand{\cA}{\mathcal{A}} \newcommand{\cB}{\mathcal{B}}
\newcommand{\cC}{\mathcal{C}} \newcommand{\cD}{\mathcal{D}}
\newcommand{\cE}{\mathcal{E}} \newcommand{\cF}{\mathcal{F}}
\newcommand{\cG}{\mathcal{G}} \newcommand{\cH}{\mathcal{H}}
\newcommand{\cI}{\mathcal{I}} \newcommand{\cJ}{\mathcal{J}}
\newcommand{\cK}{\mathcal{K}} \newcommand{\cL}{\mathcal{L}}
\newcommand{\cM}{\mathcal{M}} \newcommand{\cN}{\mathcal{N}}
\newcommand{\cO}{\mathcal{O}} \newcommand{\cP}{\mathcal{P}}
\newcommand{\cQ}{\mathcal{Q}} \newcommand{\cR}{\mathcal{R}}
\newcommand{\cS}{\mathcal{S}} \newcommand{\cT}{\mathcal{T}}
\newcommand{\cU}{\mathcal{U}} \newcommand{\cV}{\mathcal{V}}
\newcommand{\cW}{\mathcal{W}} \newcommand{\cX}{\mathcal{X}}
\newcommand{\cY}{\mathcal{Y}} \newcommand{\cZ}{\mathcal{Z}}

\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}

\newenvironment{proof}{{\bf Proof:  }}{\hfill\rule{2mm}{2mm}}
\newenvironment{proofof}[1]{{\bf Proof of #1:  }}{\hfill\rule{2mm}{2mm}}
\newenvironment{proofofnobox}[1]{{\bf#1:  }}{}\newenvironment{example}{{\bf Example:  }}{\hfill\rule{2mm}{2mm}}

%\renewcommand{\thesection}{\lecnum.\arabic{section}}
%\renewcommand{\theequation}{\thesection.\arabic{equation}}
%\renewcommand{\thefigure}{\thesection.\arabic{figure}}

\newtheorem{fact}{Fact}[section]
\newtheorem{lemma}[fact]{Lemma}
\newtheorem{theorem}[fact]{Theorem}
\newtheorem{definition}[fact]{Definition}
\newtheorem{corollary}[fact]{Corollary}
\newtheorem{proposition}[fact]{Proposition}
\newtheorem{claim}[fact]{Claim}
\newtheorem{exercise}[fact]{Exercise}
\newtheorem{note}[fact]{Note}
\newtheorem{conjecture}[fact]{Conjecture}

\newcommand{\size}[1]{\ensuremath{\left|#1\right|}}
\newcommand{\ceil}[1]{\ensuremath{\left\lceil#1\right\rceil}}
\newcommand{\floor}[1]{\ensuremath{\left\lfloor#1\right\rfloor}}

%END MACROS
%Page style
\pagestyle{fancy}

\listfiles

\raggedbottom

\lhead{2017-04-04}
\rhead{W. Justin Toth CO750-Approximation Algorithms Report} %CHANGE n to ASSIGNMENT NUMBER ijk TO COURSE CODE
\renewcommand{\headrulewidth}{1pt} %heading underlined
%\renewcommand{\baselinestretch}{1.2} % 1.2 line spacing for legibility (optional)

\begin{document}
\paragraph{Paper Title:} Popularity, Mixed Matchings, and Self-duality
\paragraph{Authors:} Chien-Chun Huang and Telikepalli Kavitha
\paragraph{Reference:} \cite{huang2017popularity}

\section{Preliminaries}
\paragraph{}
We will begin by introducing the definitions of the main objects of study, and some useful notation. Then we will state the theorems to be shown in the later sections of this report.
\paragraph{}
As in most problems in the stable matching family we are given a graph $G=(V,E)$ where each vertex has ranking over its neighbours. We will assume throughout that these preference orders are strict, and a vertex prefers any neighbour to being unmatched. We will denote by $N(v)$ the set of neighbours of vertex $v$, and by $\delta(v)$ the set of edges incident with $v$. We will use the notation $>_v$ to describe $v$'s preference order over $N(v)$. Precisely for $u, w \in N(v)$, we write $u >_v w$ if $u$ is preferred by $v$ to $w$. We will use $\geq_v$, $\leq_v$, $=_v$, and $<_v$ analogously.
\paragraph{}
One way to think of $\textit{popular}$ matchings is as matchings which never lose an election to any other matching. Since we will speak often about partners in matchings, we introduce the notation: if $M$ is a matching then $M(u) = v$ provided $uv \in M$. For any $u \in V$ and $v,v' \in N(v)$ we define $vote_u(v,v')$ as
$$vote_u(v,v') := \begin{cases}
1, \text{ if } v >_u v' \\
-1, \text{ if } v <_u v' \\
0, \text{ otherwise (ie. } v = v').
\end{cases}$$
To decide the winner of the election we some the votes. The aggregate voting between two matchings $M$ and $M'$ is
$$\Delta(M, M') = \sum_{u \in V} vote_u(M(u), M'(u)).$$
Observe that $\Delta(M,M') \geq 0$ if at least as many vertices prefer their partner in $M$ to their partner in $M'$. A $\textit{popular}$ matching $M$ is  a matching where
$$\Delta(M, M') \geq 0$$
for any matching $M'$.
\paragraph{}
We are interested in the $\textit{max weight popular matching}$ problem in this report. That is, given a weight function $w: E \rightarrow \R_+$ find a popular matching $M$ on $G$ which maximizes $w(M) := \sum_{e \in M} w(e).$ A standard approach to such problems is to attempt to formulate them as linear programs. This will motivate our definitions of fractional popular matchings to follow. In this approach the major question is whether or not your linear programming formulation gives back integral solutions which can be used to solve the original combinatorial problem. In this paper we will prove that there is an integral formulation for a special case, and a $\frac{1}{2}$-integral formulation in general.
\paragraph{}
The fractional matching polytope $FM_{G}$ is defined as
$$FM_{G} := \{x \in \R_{\geq 0}^{E} : x(\delta(v)) \leq 1, \forall v \in V \}.$$
This object has been well-studied in combinatorial optimization. We want to study the analogous $\textit{popular fractional matching polytope}$ which we will call $P_G$. To formally define $P_G$ we will need to extend our notions related to voting as used in defining popular matchings. Let $x \in FM_G$. We define
$$vote_u(x, v') := \sum_{uv \in \delta(u)} x_{uv} vote_u(v,v') = x(\delta^{>v'}(u)) - x(\delta^{<v'}(u)),$$
where $\delta^{>v'}(u) := \{vu \in \delta(u) : v >_u v'\}$ denotes the set of edges $u$ prefers to $uv'$, and $\delta^{<v'}(u)$ is defined analogously.  We define $vote_u(v',x) = -vote_u(x,v')$. Now we have a means of voting between two fractional matchings $x,y \in FM_G$:
$$vote_u(x,y) := \sum_{uv' \in \delta(u)} y_{uv'}vote_u(x,v'), $$
and the total votes:
$$\Delta(x,y) := \sum_{u \in V} vote_u(x,y).$$
Now we can formally describe $P_G$ as 
$$P_G = \{ x\in FM_G: \Delta(x,y) \geq 0, \forall y \in FM_G\}.$$
\paragraph{}
We finally have sufficient definitions to describe the results of this paper. In section \ref{sec:formulation} we describe a useful extended formulation which describes $P_G$ that will come into play when studying popular fractional matchings throughout this work. In section \ref{sec:special} we build up the following theorem
\begin{theorem}\label{th:special}
Let $G = (A \cup B, E)$ be a bipartite graph with strict preferences. If $G$ admits a perfect stable matching the $P_G$ is integral.
\end{theorem}
Building upon this theorem we show that $\frac{1}{2}$-integrality holds in general in section \ref{sec:general}.
\begin{theorem}\label{th:general}
Let $G=(V,E)$ be an arbitrary graph with strict preferences. Then the extreme points of $P_G$ lie in $\{0,\frac{1}{2}, 1\}^E$.
\end{theorem}
In section \ref{sec:complexity} we discuss some results showing that max weight popular matching is $NP$-hard and under the Unique Games Conjecture is difficult to approximate. We conclude with a section discussing open problems.
\section{Extended Formulation}\label{sec:formulation}

\paragraph{}
Here we present the extended formulation $P'_G$ which we will use to study the popular fractional matching polytope. The properties of $P'_G$ were first studied in \cite{kavitha2011popular}. For a more rigorous treatment of its connection to $P_G$ refer to that paper. Here we try to present ideas about how $P'_G$ was derived.

\paragraph{}
We begin by augmenting our graph $G$ with a set of last resort neighbours $\ell(u)$ for every $u \in V$. The vertices $\ell(u)$ are only connected to their respective vertex $u$. The preference lists of each $u$ are modified so that $\ell(u)$ is their last choice. That is for all $v \in N(u)$, $v \geq_u \ell(u)$. The vertex set and edge set now look like
$$V(G) = V \cup \bigcup_{u \in V} \ell(u) \quad\text{and}\quad E(G) = E \cup \bigcup_{u \in V} u\ell(u).$$  With the addition of the last resort vertices we may now assume that every $v \in V$ (the original vertices) is fully matched in any popular fractional matching.

\paragraph{}
Fix a fractional matching $x \in FM_G$.  We seek a linear program which characterizes when $x$ is popular. Let $w: E(G) \rightarrow \R_+$ be the weight function given by assigning  $$w(ab) := vote_a(b,x) + vote_b(a,x) = x(\delta^{<b}(a)) - x(\delta^{>b}(a)) + x(\delta^{<a}(b)) - x(\delta^{>a}(b))$$
for any $ab \in E$, and assigning
$$w(u\ell(u)) := vote_u(\ell(u),x) = -x(\delta^{>\ell(u)}(u)).$$
Observe that if $y \in FM_G$ fully matches each $u \in V$ then $\sum_{ab \in E(G)} w(ab)y_{ab} = \Delta(y,x).$ So the fractional matching $x$ is popular if and only if the optimal value of the following LP is $0$ (it cannot be less than $0$ as $x$ is a feasible solution, and $\Delta(x,x) = 0$) :
\begin{align*}
\max\  &\sum_{e \in E(G)} w(e) y_{e} \\
\text{s.t.}\ y(\delta(u)) &= 1 &\forall u \in V \\
y_e &\geq 0 &\forall e \in E(G).
\end{align*}
Consider the dual to previous linear program, to be denoted $(LP1)$:
\begin{align*}
\min\ &\sum_{u \in V} \alpha_u \\
\text{s.t.}\ \alpha_a + \alpha_b &\geq w(ab) &\forall ab \in E \\
\alpha_u &\geq w(u\ell(u)) &\forall u \in V.
\end{align*}
If $x$ is popular then the optimal value of $(LP1)$ is $0$. Also note that if $x$ is not popular then the optimal value is greater than $0$. So one means of finding popular fractional matchings would be to let $x$ vary as a fractional matching and minimize $\sum_{u} \alpha_u$. We call that linear program $(LP2)$:
\begin{align*}
\min\ &\sum_{u \in V} \alpha_u \\
\text{s.t.}\ \alpha_a + \alpha_b &\geq w(ab) &\forall ab \in E \\
\alpha_u &\geq w(u\ell(u)) &\forall u \in V \\
x(\delta(u)) &= 1 &\forall u \in V \\
x_e &\geq 0 &\forall e \in E(G).
\end{align*}
As discussed above, the optimal value of $(LP2)$ is $0$, and so a corresponding optimal solution for $(LP2)$ is a vector $(x,\alpha) \in \R^{E(G)}\times \R^{V}$ where $x$ is a popular fractional matching and $\sum_{u \in V} \alpha_u = 0$. Such $\alpha$ is called a $\textit{witness}$ to the popularity of $x$. The polytope of optimal solutions of $(LP2)$ is what we will use as our extended formulation for the popular fractional matching polytope. We can describe it as $P'_G$, the set of solutions to the set of linear inequalities:
\begin{align*}
\sum_{u \in V} \alpha_u &= 0\\
\alpha_a + \alpha_b &\geq vote_a(b,x) + vote_b(a,x) &\forall ab \in E \\
\alpha_u &\geq vote_u(\ell(u),x) &\forall u \in V \\
x(\delta(u)) &= 1 &\forall u \in V \\
x_e &\geq 0 &\forall e \in E(G).
\end{align*}
Thus $P'_G$ involves $O(|E| + |V|)$ constraints, and hence provides a polynomial size description of popular fractional matchings. The following lemma demonstrates a very interesting property of $(LP2)$
\begin{lemma}\label{lemma:self-dual}
The linear program $(LP2)$ is its own dual program program. That is $(LP2)$ is self-dual.
\end{lemma}
\begin{proof}
Consider the dual to $(LP2)$ in variables $y \in \R^{E(G)}$ and $\beta \in \R^V$:
\begin{align*}
\max\ &\sum_{u \in V} \beta_u \\
\text{s.t.} \beta_a + \beta_b + y(\delta^{<b}(a))  - y(\delta^{>b}(a)) + y(\delta^{<a}(b)) - y(\delta^{>a}(b)) &\leq 0 &\forall ab \in E \\
\beta_u - y(\delta^{>\ell(u)}(u)) &\leq 0 &\forall u \in V \\
y(\delta(u)) &= 1 &\forall u \in V \\
y_e &\geq 0 &\forall e \in E(G).
\end{align*}
If we perform the substitutions $\alpha_u = -\beta_u$ for each $u \in V$ and $x_e = y_e$ for each $e \in E(G)$ then the above dual LP becomes exactly the same as $(LP2)$.
\end{proof}
\paragraph{}
This lemma leads to the following very useful lemma, which follows immediately from the previous result via complementary slackness.
\begin{lemma}\label{lemma:self-dual-cs}
Let $(x, \alpha^x) \in P'_G$.  For every $ab \in E$, if $x_{ab} > 0$ then the ``covering constraint" in $P'_G$ for $ab$ is tight. Formally we have
$$\alpha_a^x + \alpha_b^x = x(\delta^{<b}(a)) - x(\delta^{>b}(a)) + x(\delta^{<a}(b)) - x(\delta^{>a}(b)).$$
\end{lemma}
\section{Integrality of $P_G$ in a special case}\label{sec:special}
\paragraph{}
In this section we turn our attention to theorem \ref{th:special}. So suppose that our graph $G$ is bipartite with bipartition $A \cup B$. Also suppose that $G$ admits a perfect stable matching. This condition actually implies that every popular matching is perfect. This follows since stable matchings always match the same set of vertices, and stable matchings are minimum size popular matchings \cite{huang2011popular}. It is not hard to see that this extends to popular fractional matchings, meaning every popular fractional matching fully matches each $u \in V$ to its true neighbours (not last resort vertices). 
\paragraph{}
Let $(x, \alpha^x) \in P'_G$. We want to write $x$ as a convex combination of popular matchings. First we want bounds on the values the witness vector $\alpha^x$ takes
\begin{lemma}
For every $u \in V$, $\alpha_u^x \in [-1,1]$.
\end{lemma}
\begin{proof}First
$$\alpha^x_u \geq vote_u(\ell(u), x) = -1 \quad\text{(since }x_{u,\ell(u)} = 0\text{).}$$
Now let $v$ be the least preferred neighbour of $u$ such that $x_{uv} >0$. Then
$$vote_u(v,x) = -(1-x_{uv}).$$
Also
$$vote_v(u,x) = x(\delta^{<u}(v))-x(\delta^{>u}(v))= 1-x_{uv}.$$
So by lemma \ref{lemma:self-dual-cs},
$$\alpha_u + \alpha_v = vote_u(v,x) + vote_v(u,x) = -(1-x_{uv}) + (1-x_{uv}) = 0.$$
Since $\alpha_v \geq -1$ this implies $\alpha_u \leq 1$.
\end{proof}
\paragraph{}
Thinking ahead, we want to write $x$ as a convex combination of popular matchings. In doing this we will use a witness vector in $\{-1,1\}^V$ to certify the popularity of each chosen matching. From our previous lemma we can think of each $\alpha^x_u$ as a convex combination of $-1$ and $1$. For each $a \in A$ let $r_a$ denote the fraction of popular matchings in the convex combination forming $x$ whose $\alpha_a^x$ value is assigned $1$. Similarly define for each $b \in B$ the value $r_b$ to be the fraction assigning $\alpha^x_b$ to $-1$. Precisely we have $2r_a -1 = \alpha^x_a$ and $1-2r_b = \alpha^x_b$.
\paragraph{}
Let $X_a$ be the array for the assignment of $a\in A$ in $x$. Each cell of $X_a$ corresponds to some $b \in N(a)$ and the length of the cell is $x_{ab} > 0$ (we omit $0$-length cells). The cells in $X_a$ are arranged in increasing order with respect to $a$'s preferences. The first $r_a$ fraction of $X_a$ is called the $\textit{positive}$ subarray and the remaing the $\textit{negative}$ subarray. We split and reorder the arrays as shown in the following figure to obtain $X'_a$
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{Xa}
\end{figure}
We do the same thing for each $b \in B$, creating $X_b$. But the only difference is that we arrange $X_b$'s cell in decreasing order of $b$'s preferences. Also notice that since $r_b$ corresponds to the negative $\alpha_b$ cells, the array begins with the negative subarray. We again split and reorder to obtain $X'_b$ as shown
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{Xb}
\end{figure}
One interesting thing about our choice of the $r_a$'s and $r_b$'s is that they allow us to rewrite our covering constraints for $ab$ in a simpler form.
\begin{lemma}\label{lemma:covering}
For any $ab \in E$ we have
$$x_{ab} + x(\delta^{<b}(a)) + x(\delta^{<a}(b)) \leq r_a + (1-r_b),$$
and if $x_{ab} > 0$ then the above is tight.
\end{lemma}
\begin{proof}
Take the constraint
$$ \alpha^x_a + \alpha^x_b \geq x(\delta^{<b}(a)) - x(\delta^{>b}(a)) + x(\delta^{<a}(b)) - x(\delta^{>a}(b)) $$
and rewrite using $\alpha^x_a = 2r_a-1$ and $\alpha^x_b = 1-2r_b$. Further substitute $-x(\delta^{>b}(a)) = -(1 -x_{ab} - x(\delta^{<b}(a)))$ and $-x(\delta^{>a}(b)) = -(1- x_{ab} - x(\delta^{<a}(b)))$. After simplifying the inequality in the lemma statement is obtained. By lemma \ref{lemma:self-dual-cs} this inequality is tight when $x_{ab} > 0$.
\end{proof}

\paragraph{}
Now we show how to obtain popular matchings whose convex hull contains $x$. Form a table $T$ by stacking each $X'_u$ for $u \in V$ in rows. The width of $T$ is $1$, so for any $t \in [0,1)$ we can define a set of edges $M_t$ as follows:
\begin{enumerate}
\item Form a vertical line at distance $t$ from the left end of $T$
\item For each $u \in V$ the vertical line intersects (or touches the left boundary of) a cell in $X'_u$. We will denote this cell by $c_u(t)$.
\item Form the set $M_t$ as $M_t := \{uv : u \in V \text{ and } uv \text{ corresponds to cell } c_u(t)\}$.
\end{enumerate}
\paragraph{}
We need to verify two things: the first is that $M_t$ is a matching, and the second is that $M_t$ is indeed popular. This will be done in the following two subsections. Supposing we are successful in the promised arguments to follow, we will show how this allows us to find a set of popular matchings for which $x$ is a convex combination. Execute the following algorithm:
\begin{enumerate}
\item Initialize $i=0$.
\item Sweep a vertical line from left to right across $T$. Denote its distance from the left by $t$. 
\item Whenever a new cell is encountered fix matching $M^i = M_t$ and fix $t_i = t$. 
\end{enumerate}
Say we find $k$ such matchings. We can then write $x$ as:
$$x = t_1\chi(M^0) + (t_2 - t_1)\chi(M^1) + \dots + (1 - t_{k-1}) \chi(M^{k-1})$$
where $\chi(\cdot)$ if the function mapping sets to incidence vectors.

\subsection{$M_t$ is a matching}
\paragraph{}
Let $ab$ be an edge with $x_{ab} > 0$. To show that $M_t$ is a matching it suffices to show that the cell corresponding to $ab$ is perfectly aligned in its position in $X'_a$ with its position in $X'_b$. Then when a vertical line is dropped through the $ab$-cell in $X'_a$ it also dropped through the $ab$-cell in $X'_b$. Thus each vertex is matched to exactly one partner in $M_t$.

\begin{proof} There are two cases to consider: When $x(\delta^{<b}(a)) \geq r_a$ (the case $x(\delta^{<a}(b)) \geq 1-r_b$ is absolutely symmetric) and the case when both $x(\delta^{<b}(a)) < r_a$ and $x(\delta^{<a}(b)) < 1-r_b$.
\paragraph{Case 1}
Suppose that $x(\delta^{<b}(a)) \geq r_a$. Then $x_{ab}$ is in $a$'s negative subarray. Let $d = x(\delta^{<b}(a)) - r_a$ be the distance from the left $X'_a$ to $x_{ab}$. Let $d' = x(\delta^{<a}(b))$ be the distance to $x_{ab}$ from the right of the positive subarray of $X'_b$ (see the figure). From lemma \ref{lemma:covering} we have
$$d + x_{ab} + d' = 1-r_b.$$
Thus $x_{ab}$ is perfectly aligned as desired. Further since $d \geq 0$, $x_{ab}$ lies in $X'_b$'s positive subarray.
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{geqRa}
\caption{Case 1}
\end{figure}
\paragraph{Case 2}
Suppose $x(\delta^{<b}(a)) < r_a$ and $x(\delta^{<a}(b)) < 1-r_b$. In this case we have separated the cell corresponding to $x_{ab}$. We will describe the proof below, but as before it is more clear from the figure
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{lessRa}
\caption{Case 2}
\end{figure}
Let $d_0 = x(\delta^{<b}(a))$ be the distance from $X'_a$'s positive subarray to $x_{ab}$. Let $d_1 = x(\delta^{<a}(b))$ be the distance from the right of $X'_b$'s positive subarray to the right of $x_{ab}$. Let $x^0_{ab} = r_a-d_0$ be the length of the portion of the $x_{ab}$ cell on the right end of $X'_a$. Let $x^1_{ab} = 1-r_b - d_1$ be the length of the portion of the $x_{ab}$ cell on the left end of $X'_b$. By lemma \ref{lemma:covering} we know that
$$x_{ab} + d_0 + d_1 = r_a + (1-r_b)$$
and thus $x_{ab} = x^0_{ab} + x^1_{ab}$. Therefore as in the previous case the cells for $x_{ab}$ are perfectly aligned. Again observe that the cell portions in the positive subarray of $X'_a$ are aligned in the negative subarray of $X'_b$, and vice versa. This additional observation will be useful when we move to popularity.
\end{proof}
\subsection{$M_t$ is popular}
\paragraph{}
Our plan is to exhibit a witness vector $\alpha \in \{-1,1\}^V$ such that $(\chi(M_t), \alpha) \in P'_G$. Then $M_t$ is popular. We construct $\alpha$ as follows, for every $u \in A\cup B$: set $\alpha_u = -1$  if the vertical line intersect the negative subarray of $X'_u$, else set $\alpha_u = 1$ (the vertical line intersects or touches the boundary of the positive subarray).
\paragraph{}
Let $ab \in M_t$. Then, as an artifact of the proof that $M_t$ is a matching, the vertical line intersects the positive subarray of $X'_a$ if and only if it doesn't intersect the positive subarray for $X'_b$. Thus $\alpha_a + \alpha_b = 0$. Since $M_t$ is perfect this implies that
$$\sum_{u \in A\cup B} \alpha_u = 0.$$
Since $\alpha_u \geq -1 = vote_u(\ell(u), M_t)$ as $M_t$ is perfect, it only remains to verify the covering constraints. That is for all $ab \in E$
$$\alpha_a + \alpha_b \geq vote_a(b,M_t) + vote_b(a,M_t).$$

\begin{proof} Recall that we use $c_u(t)$ to denote the cell of $X'_u$ which the vertical line at distance $t$ intersects. Then the cell $c_u(t)$ has a correspondence with $uM_t(u)$. We delineate the following three cases: $1)$ both $c_a(t)$ and $c_b(t)$ are lie in their respective positive subarrays, $2)$ one of $c_a(t)$ or $c_b(t)$ is positive and the other negative, $3)$ or both $c_a(t)$ and $c_b(t)$ are negative. 
\paragraph{Case 1}
Suppose that both $c_a(t)$ is positive and $c_b(t)$ is positive. Then $\alpha_a = \alpha_b = 1$. Since $vote_a(b,M_t) + vote_b(a,M_t) \leq 2$ this constraint is easily satisfied.
\paragraph{Case 2}
Suppose that one of $c_a(t), c_b(t)$ is positive and the other is negative. Then $\alpha_a + \alpha_b = 0$. From lemma \ref{lemma:covering} we know
$$x(\delta^{<b}(a)) + x_{ab} + x(\delta^{<a}(b)) \leq r_a + (1-r_b)$$
Hence for $a$ and $b$ to be matched to worse partners than each other, we need $c_t(a)$ and $c_t(b)$ to be in their respective positive subarrays. That cannot happen in this case, so either $a$ and $b$ matched, or at least one of them prefers the other to their partner. Hence 
$$vote_a(b,M_t) + vote_b(a,M_t) \geq 0$$
as desired.
\paragraph{Case 3}
Suppose that both $c_a(t)$ and $c_b(t)$ are negative. Then $\alpha_a + \alpha_b = -2$. Again using the inequality 
$$x(\delta^{<b}(a)) + x_{ab} + x(\delta^{<a}(b)) \leq r_a + (1-r_b)$$
we see that both $a$ and $b$ are matched to better partners than each other. Therefore $vote_a(b,M_t) + vote_b(a,M_t) = -2$. This is clear from the following picture:
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{covering}
\caption{In Case 3, $c_a(t)$ and $c_b(t)$ are in the common red (negative) area for both. Since $x_{ab}$ lies in the sum length of $r_a + 1-r_b$, both $a$ and $b$ prefer their match to each other.}
\end{figure}
\end{proof}
\section{$\frac{1}{2}$-Integrality of $P_G$ in general}\label{sec:general}
\paragraph{}
We now return our attention to general, ie. non-bipartite, graphs $G = (V,E)$ and do not assume the presence of a perfect stable matching. In this case we are interested in proving theorem \ref{th:general}, which says that $P_G$ is $\frac{1}{2}$-integral in these instances. Our strategy is to map $G$ to a bipartite graph with a perfect stable matching, use the previous result, and map what we get back.
\paragraph{}
We begin by describing the bipartite graph $H$ obtained from $G$. The vertex set is $V(H) := V \cup V'$, where $V' = \{v' : v \in V\}$, and so $H$ has two vertices $v$ and $v'$ for each vertex in $G$. The edge set is $E(H) := \{uv', vu': uv \in E\} \cup \{uu' : u \in V(G)\}$. Hence $V$, $V'$ forms a bipartition of $V(H)$. Now we describe the preference lists. For each $v \in V$, if $u >_v w$ then $u'>_v w'$ in $H$ and $u >_{v'} w$ in $H$. In $H$ each vertices last choice is their ``twin" vertex. So $u' \geq_v v'$ for all $u' \in N(v)$, and similarly $u \geq_{v'} v$ for all $u \in N(v')$.
\begin{lemma}
$H$ admits a stable matching.
\end{lemma}
\begin{proof}
Let $S$ be a stable matching in $H$. Suppose without loss of generality that $S$ exposes $u \in V(H)$. Let $S' := \{uv' :u'v \in S\} \cup \{u'v: uv' \in S\}$ be the ``twin" matching of $S$. By symmetry of the preference orders, since $S$ is stable so is $S'$. By our construction of $S'$, $S'$ exposes $u'$. But further since each stable matching matches the same vertex set, this implies that $S$ also exposes $u'$. But then the edge $uu'$ blocks $S$, a contradiction.
\end{proof}

\paragraph{}
Thus by theorem \ref{th:special}, $P_H$ is integral. Now we describe the mapping alluded to an the outset of this section. Let $f: P_G \rightarrow P_H$ be defined as follows.
$$f(x) = z \text{ where } z_{uv'}=z_{u'v} =x_{uv} \text{ and } z_{uu'} = x_{u\ell(u)}.$$
We need to verify that $f$ indeed maps popular fractional matchings to popular fractional matchings. It is clear that $z$ is a fractional matching on $H$. We need to check that $z$ is popular.
\begin{lemma}
Let $x \in P_G$. Then $z = f(x) \in P_H$.
\end{lemma}
\begin{proof}
We construct a witness $\beta$ to the popularity of $z$. Let $\alpha$ be a witness to the popularity of $x$. For each $u \in V$ set $\beta_u = \beta_{u'} = \alpha_u$. Now $\sum_{v\in V(H)} \beta_v = \sum_{v \in V} \beta_v + \sum_{v' \in V'} \beta_v' = 2 \sum_{u \in V} \alpha_u = 0$. Now we need to check the covering constraints. Let $ab \in E(G)$. Since $(x,\alpha)$ satisfies the $ab$-covering constraint for $P'_G$, it immediate that $(z,\beta)$ satisfies the covering constraints for $ab'$ and $a'b$ in $P'_H$.
\paragraph{}
It remains to consider covering constraints for $uu' \in E(H)$. Consider $u\ell(u) \in E(G)$. We have $\alpha_u \geq x_{u\ell(u)} - 1$ from the description of $P'_G$. Now observe that the right-hand side of the covering constraint for $uu'$ in $P'_H$ is
$$vote_u(u', z) + vote_u'(u,z) = 2(x_{u\ell}-1).$$
The left-hand side of the covering constraint for $uu'$ in $P'_H$ is
$$\beta_u + \beta_u' = 2\alpha_u \geq 2(x_{u\ell(u)}-1)$$
and hence the covering constraint for $uu'$ is satisfied. It is trivial to see that the last-resort covering constraints in $P'_H$ are satisfied. Hence $z \in P_H$.
\end{proof}
\paragraph{}
Now we need a mapping that will take us back from $P_H$ to $P_G$. We want this mapping to be the inverse of $f$ when restricted to fractional matchings in the image of $f$. Let $h: P_H \rightarrow P_G$. We want that $h \circ f (x) = x$ for all $x \in P_G$. To satisfy this we define $h$ as follows:
$$h(z) = y \text{ where } y_{uv} = (z_{uv'} + z_{u'v})/2 \text{ and } y_{u\ell(u)} = z_{uu'}.$$
Verifying that $h(z)$ is popular is a routine calculation as in the previous lemma. The witness vector chosen is  $\alpha$ where $\alpha_u = (\beta_u + \beta_u')/2$ and $\beta$ is a witness to the popularity of $z$.
\paragraph{}
Now we can show how to write $x \in P_G$ as a convex combination of $\frac{1}{2}$-integral popular matchings. Let $z=f(x)$. Then $z \in P_H$. By theorem \ref{th:special}, that there exist $M_1, \dots, M_k$ and $\lambda \in \R_+^k$ such that $\sum_{i=}^k \lambda_i = 1$ and 
$$z = \sum_{i=1}^k \lambda_i \chi(M_i).$$
Now we have $x = h(z)$, and from the definition of $h$, $h$ distributes over scalar multiplication so we have
$$x = h(z) = \sum_{i=1}^k \lambda_i h(\chi(M_i)).$$
Now from the definition of $h$, $h(\chi(M_i)) \in \{0,\frac{1}{2}, 1\}^{E(G)}$, so we have shown that $x$ is a convex combination of $\frac{1}{2}$-integral popular matchings.$\blacksquare$
\section{Complexity Results}\label{sec:complexity}
\paragraph{}
The previous section showed that we can find max-weight popular $\frac{1}{2}$-integral matchings in polynomial time. This section seeks to study the hardness of max-weight popular matching. We will show that not only is this problem $NP$-hard, but hard to approximate under the Unique Games Conjecture.
\paragraph{}
We give a reduction from VERTEX COVER $(VC)$. Let $H = (V_H, E_H)$ be an instance of $(VC)$ with $|V(H)| = n$. From $H$ we will construct a graph $G=(V,E)$ with preferences, and a weight function $w: E \rightarrow \{1,2\}$. The graph $G$ will be the complete graph on $4n$ vertices. For each $v \in V(H)$ we have $4$ vertices $v_0, v_1, v_2, v_3 \in V(G)$. Below we present the preference lists of $v_0,v_1,v_2,v_3$. In these lists we use $\pi(u_0: u \in N(v))$ to denote an arbitrary permutation on  $\{u_0: u \in N(v) \text{ with respect to $H$}\}$, and we use $\pi(\dots)$ to denote an arbitrary permutation of the vertices not mentioned so far in the particular list. We use these notations to indicate that our reductions are indifferent to how these parts of the preference lists are arranged.
\begin{align*}
v_0 &: v_1 > \pi(u_0: u \in N(v)) > v_2 > v_3 >\pi(\dots) \\
v_1 &: v_0 > v_2 > v_3 > \pi(\dots) \\
v_2 &: v_1 > v_0 > v_3 > \pi(\dots) \\
v_3 &: v_1 > v_2 > v_0 > \pi(\dots).
\end{align*}
The weight function $w$ is described by $w(v_0v_2) = w(v_1v_3) = 2$ for each $v \in V(H)$ and $w(e) = 1$ for all other edges $e\in E(G)$. 
\paragraph{}
In the reductions we rely critically on the following characterization of popular matchings from \cite{huang2011popular}.
\begin{lemma}\label{lemma:char}
\cite{huang2011popular} Let $M$ be a perfect matching in $G$. For all $uv \in E\backslash M$ label $uv$ with $(vote_u(v,M(u)), vote_v(u,M(v)))$. Then $(1,1)$ edges are blocking edges in the stable matching sense. Let $G_M$ be the graph obtained from $G$ by deleting $(-1,-1)$ edges (doesn't hurt popularity). Then $M$ is popular if and only if in $G_M$:
\begin{enumerate}
\item There is no $M$-alternating cycle containing a blocking edge, and
\item there is not $M$-alternating path with two or more blocking edges.
\end{enumerate}
\end{lemma}

\begin{lemma}\label{lemma:firstreduc}
Let $C$ be a vertex cover of size $k$ in $H$. Then there exists a popular matching $M$ of $G$ with $w(M) = 4n-2k$. This gives one direction of the desired reduction.
\end{lemma}
\begin{proof}
Construct matching $M$ by adding two edges for each $v \in V(H)$:
$$\text{ if $v \in C$ then add $v_0v_1, v_2v_3$ to $M$, otherwise add $v_0v_2, v_1v_3$ to $M$}.$$
Then $w(M) = (1+1)k + (2+2)(n-k) = 4n - 2k$ as desired. We need only verify popularity of $M$. This is done through checking the two conditions of lemma \ref{lemma:char} on $G_M$. One can observe there are four classes of edges in $G_M$:
\begin{enumerate}
\item Edges of $M$.
\item Edges $v_0v_2, v_1v_2, v_1v_3$ for all $v \in C$. The label on these edges is $(-1,1)$.
\item Edges $v_0v_1, v_1v_2$ for all $v \not\in C$. The label on these edges is $(1,1)$, that is they are blocking edges.
\item The edge $v_0u_0$ where $vu \in E(H)$ and exactly one of $u,v \in C$. The label on this edge is either $(1,-1)$ or $(-1,1)$.
\end{enumerate}
It is not hard to verify the first two conditions of lemma \ref{lemma:char} from this point. To see the alternating cycle condition, one observes that the edges containing any $v_3$ are either in $M$ or $(1,1)$ edges, hence the degree of $v_3$ is one. Since blocking edges are either $v_0v_1$ of $v_1v_2$ and their partner is $v_3$ there cannot be an alternating cycle containing a blocking edge.
\paragraph{}
Suppose $P$ is an $M$-alternating path containing either $v_0v_1$ or $v_1v_2$ for $v \not\in C$. It cannot contain both as there is no such alternating path in $G_M$. Then the path $P$ needs to use a class $4$ edge $v_0u_0$ to get out of the ``$v$" edges and find another blocking edge. Since $v \not\in C$, such $u$ is in $C$. Thus $M(u_0) = u_1$. Now the neighbours of $u_1$ are $u_2$ and $u_3$ only. Since $u_2$ and $u_3$ are not incident to any blocking edges, $P$ cannot find a second blocking edge.
\end{proof}
\begin{lemma}\label{lemma:6.2}
Let $M$ be a popular matching in $G$. Then there does not exist $v_su_j \in M$ with $v\neq u$ and $s,t \in \{0,1,2,3\}$.
\end{lemma}
\begin{proof}
The proof is by case analysis using the characterization from lemma \ref{lemma:char}. They show that allowing such an edge contradicts the characterization. The computations are routine in popular matchings, so we'll enumerate the cases and spare the details. The cases are
\begin{enumerate}
\item $s,t \in \{0,1\}$
\item $s, \in \{0,1\}$ and $t=2$
\item $s \in \{0,1\}$ and $t=3$
\item $s,t \in\{2,3\}$.
\end{enumerate}
\end{proof}
\begin{lemma}\label{lemma:6.3}
Suppose $M$ is a popular matching in $G$. Then 
\begin{enumerate}
\item For all $v_0,v_1,v_2,v_3 \in V(G)$, either $v_0v_1,v_2v_3 \in M$ or $v_0v_2v_1v_3 \in M$
\item If $vu \in E(H)$ then either $v_0v_1 v_2v_3 \in M$ or $u_0u_1, u_2u_3 \in M$.
\end{enumerate}
\end{lemma}
\begin{proof}
The previous lemma (lemma \ref{lemma:6.2}) tells us that $v_0,v_1,v_2,v_3$ are necessarily matched together. But if $v_0v_3, v_1v_2 \in M$ then $v_0,v_3,v_2,v_1,v_0$ forms an $M$-alternating cycle in $G_M$ containing blocking edge $v_0v_1$. Hence the options stated in $1.$ are the only possibilities. For the second part one can similarly find an alternating path containing two blocking edges.
\end{proof}
\begin{lemma}\label{lemma:secondreduc}
If $M$ is a popular matching in $G$ with $w(M) \geq 4n-2k$ then there is a vertex cover $C$ in $H$ with $|C| \leq k$. This gives the other direction of the desired reduction.
\end{lemma}
\begin{proof}
From part $1)$ of lemma \ref{lemma:6.3}, for all $v \in V(H)$ either $v_0v_1, v_2v_3 \in M$ or $v_0v_2,v_1v_3 \in M$. Define $C$ as
$$C := \{ v \in V(H): v_0v_1, v_2v_3 \in M\}.$$
Lemma \ref{lemma:6.3} part $2)$ implies that $H$ is a vertex cover. Now we have the equation
$$w(M) = (1+1)|C| + (2+2)(n-|C|).$$
Rearranging, using $w(M) \geq 4n-2k$,  we obtain that $|C| \leq k$.
\end{proof}
\paragraph{}
Thus we have shown the following result. Since we used a stricter version of max-weight popular matching we have a stronger $NP$-hardness result
\begin{theorem}
The max-weight popular matching problem on graph $G$ is $NP$-hard, even in the special case that $w: E(G)\rightarrow \{1,2\}$, preferences are strict, $G$ is a complete graph, and $G$ admits a stable matching.
\end{theorem}
\paragraph{}
We can obtain an inapproximability result by slightly modifying our reduction from $H$ to $G$. We change the weight function $w$ so that $w(v_0v_2) = w(v_1v_3) = 1$ for all $v \in V(H)$, and $w(e) = 0$ for all other edges of $G$. Now our previous lemmas of this section still hold, but the size-weight connection is slightly different: there is a vertex cover of $H$ of size $\leq k$ if and only if there is a popular matching of $G$ of weight $\geq 2n-2k$ (and this vertex cover can be found in polynomial time).
\begin{theorem}
There does not exist an $O(1)$-approximation algorithm for max-weight popular matching with non-negative weights under Unique Games Conjecture.
\end{theorem}
\begin{proof}
Let $C^*$ be an optimal vertex cover on $H$. Since $H$ is hardest to approximate when it contains a perfect matching, we may assume it does so. Thus $|C^*| \geq \frac{n}{2}$. Suppose for a contradiction that there exists an $\epsilon$-approximation algorithm $\cA$ for max-weight popular matching on $G$ where $\epsilon  \in (0,1)$. Compute a popular matching $M$ on $G$ using $\cA$. We split into two cases on $|C^*|$.
\paragraph{Case 1}
Suppose that $|C^*| \leq (1+\epsilon)\frac{n}{2}$. In this case $w(M) \geq \epsilon(1-\epsilon) n$. Using our reduction we can find a vertex cover $C$ with $|C| \leq (2-\epsilon+\epsilon^2)\frac{n}{2}$. Since $\frac{n}{2} \leq |C^*|$ this implies
$$|C| \leq (2-\epsilon +\epsilon^2)|C^*|.$$
\paragraph{Case 2}
In this case we obtain $C$ as above. The approximation factor is $\frac{n}{(1+\epsilon)n/2}$ which is at most $2(1-\epsilon + \epsilon^2)$.
\paragraph{}
So we can find a vertex cover $C$ with $|C| \leq (2-\epsilon + \epsilon^2)|C^*|$ in either case. Since $\epsilon \in (0,1)$ this contradiction Unique Games Conjecture.
\end{proof}
\section{Conclusions}
\paragraph{}
We have shown that given a bipartite graph with strict preferences, if the graph admits a perfect stable matching then we can solve the max-weight popular matching problem in polynomial time using linear programming. We have also shown that for arbitrary graphs we can find a max-weight $\frac{1}{2}$-integral popular matching in polynomial time. In the bipartite case, since the fractional matching polytope is integral, this implies we can actually find a mixed matching (ie a convex combination of matchings) using at most two matchings.
\paragraph{}
We have demonstrated hardness results saying that on general graphs the max-weight popular matching problem is not only $NP$-hard but also inapproximable unless Unique Games Conjecture does not hold.
\paragraph{}
The major open problems here are two: $1)$ settle the complexity of the max-weight popular matching problem on bipartite graphs, and $2)$ to resolve the popular matching problem on general graphs. The latter problem is simply the decision problem resolving existence of a popular matching on graphs with strict preferences.
\bibliography{references}
\bibliographystyle{plain}
\end{document}