### 2017-12-07 CO681-QuantumInformationProcessing/project/introduction.tex
Project start

\section{Introduction}
\paragraph{}
In this paper we explore the ways in which quantum computing can be used to improve upon the best-known classical algorithms for a wealth of fundamental problems in combinatorial optimization. In Section \ref{sec:quantum-primitives} we will discuss the various quantum procedures which appear at the heart of the otherwise classical algorithms we present. Chief among these procedures is Grover's Search \cite{grover1996fast}. In the sections which follow we discuss quantum algorithms for matchings, network flows, shortest paths, and minimum spanning trees.
\subsection{Two Models of Graphs}