### 2017-12-07 CO681-QuantumInformationProcessing/project/matchings.tex

\section{Matchings}
\paragraph{}
In a {\it maximum matching problem} we are given a graph $G=(V,E)$ and asked for find  $M\subseteq E$, maximizing $|M|$, such that every vertex of $(V,M)$ has degree at most $1$. \subsection{Bipartite Matching}
\paragraph{}
A {\it bipartite graph} is a graph where there exists a partition $A\cup B = V$ such that no vertices in $A$ are adjacent, and no vertices in $B$ are adjacent. A well known equivalent definition of a bipartite graph is one with no odd cycles. In bipartite graphs the best known classical algorithm takes time $O(n^\omega)$ \cite{mucha2004maximum} where $2\leq\omega\leq 2.4$ is the matrix multiplication constant and $n = |V|$, with the caveat that if $\omega = 2$ then algorithm a $\log n$ factor appears in the runtime.  We model our quantum algorithm on the following $O(n^{5/2})$-time classical algorithm \cite{hopkroft1973n5}:
\begin{enumerate}
\item
\end{enumerate}

\subsection{General Matching}

\subsection{Quantum Adversary}