### 2017-12-07 CO681-QuantumInformationProcessing/project/quantum-primitives.tex

\section{Quantum Primitives}\label{sec:quantum-primitives}

\subsection{Grover's Search}

\subsection{Layered Subgraphs}

\subsection{Minima Finding}