### 2018-01-17 CO749-GraphColoring/assignments/A1.tex
Wrote up problem 1

\documentclass[letterpaper,12pt,oneside,onecolumn]{article}
\usepackage[margin=1in, bottom=1in, top=1in]{geometry} %1 inch margins
\usepackage{amsmath, amssymb, amstext}
\usepackage{fancyhdr}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{theorem}
\usepackage{tikz}
\usepackage{tkz-berge}
\usepackage[braket, qm]{qcircuit}
\usepackage{hyperref}

%Macros
\newcommand{\A}{\mathbb{A}} \newcommand{\C}{\mathbb{C}}
\newcommand{\D}{\mathbb{D}} \newcommand{\F}{\mathbb{F}}
\newcommand{\N}{\mathbb{N}} \newcommand{\R}{\mathbb{R}}
\newcommand{\T}{\mathbb{T}} \newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}

\newcommand{\cA}{\mathcal{A}} \newcommand{\cB}{\mathcal{B}}
\newcommand{\cC}{\mathcal{C}} \newcommand{\cD}{\mathcal{D}}
\newcommand{\cE}{\mathcal{E}} \newcommand{\cF}{\mathcal{F}}
\newcommand{\cG}{\mathcal{G}} \newcommand{\cH}{\mathcal{H}}
\newcommand{\cI}{\mathcal{I}} \newcommand{\cJ}{\mathcal{J}}
\newcommand{\cK}{\mathcal{K}} \newcommand{\cL}{\mathcal{L}}
\newcommand{\cM}{\mathcal{M}} \newcommand{\cN}{\mathcal{N}}
\newcommand{\cO}{\mathcal{O}} \newcommand{\cP}{\mathcal{P}}
\newcommand{\cQ}{\mathcal{Q}} \newcommand{\cR}{\mathcal{R}}
\newcommand{\cS}{\mathcal{S}} \newcommand{\cT}{\mathcal{T}}
\newcommand{\cU}{\mathcal{U}} \newcommand{\cV}{\mathcal{V}}
\newcommand{\cW}{\mathcal{W}} \newcommand{\cX}{\mathcal{X}}
\newcommand{\cY}{\mathcal{Y}} \newcommand{\cZ}{\mathcal{Z}}

\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}

\newenvironment{proof}{{\bf Proof:  }}{\hfill\rule{2mm}{2mm}}
\newenvironment{proofof}[1]{{\bf Proof of #1:  }}{\hfill\rule{2mm}{2mm}}
\newenvironment{proofofnobox}[1]{{\bf#1:  }}{}\newenvironment{example}{{\bf Example:  }}{\hfill\rule{2mm}{2mm}}

%\renewcommand{\thesection}{\lecnum.\arabic{section}}
%\renewcommand{\theequation}{\thesection.\arabic{equation}}
%\renewcommand{\thefigure}{\thesection.\arabic{figure}}

\newtheorem{fact}{Fact}[section]
\newtheorem{lemma}[fact]{Lemma}
\newtheorem{theorem}[fact]{Theorem}
\newtheorem{definition}[fact]{Definition}
\newtheorem{corollary}[fact]{Corollary}
\newtheorem{proposition}[fact]{Proposition}
\newtheorem{claim}[fact]{Claim}
\newtheorem{exercise}[fact]{Exercise}
\newtheorem{note}[fact]{Note}
\newtheorem{conjecture}[fact]{Conjecture}

\newcommand{\size}[1]{\ensuremath{\left|#1\right|}}
\newcommand{\ceil}[1]{\ensuremath{\left\lceil#1\right\rceil}}
\newcommand{\floor}[1]{\ensuremath{\left\lfloor#1\right\rfloor}}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
\DeclarePairedDelimiter\norm{\lVert}{\rVert}%
%END MACROS
%Page style
\pagestyle{fancy}

\listfiles

\raggedbottom

\lhead{CO 749 - \today}
\rhead{W. Justin Toth - A1} %CHANGE n to ASSIGNMENT NUMBER ijk TO COURSE CODE
\renewcommand{\headrulewidth}{1pt} %heading underlined
%\renewcommand{\baselinestretch}{1.2} % 1.2 line spacing for legibility (optional)

\begin{document}
\section{}
\paragraph{}
Let $G$ be a graph of minimum degree at least $3$ satisfying all of the following:
\begin{itemize}
	\item Every triangle contains at most one vertex of degree three.
	\item If two vertices of degree three are adjacent then they are in triangles.
	\item No two distinct triangles share a common edge.
\end{itemize}
Suppose for a contradiction that $\abs{E(G)} < 5\abs{V(G)}/3$.
\begin{claim}\label{1.1}
	For every $v \in V(G)$, if $d(v) = 3$ then $v$ has at most one neighbour of degree $3$.
\end{claim}
\begin{proof}
	Assume for a contradiction that there exist $u,v, w \in V(G)$ satisfying
	$$d(u) = d(v) = d(w) = 3$$
	and $uv, vw \in E(G)$, i.e. $u$ and $w$ are adjacent to $v$. But then, using the second property of $G$, $v$ is in a triangle. Denote this triangle by $C$. Since every triangle contains at most one vertex of degree three, $u$ and $w$ are not contained in $C$. But then since $d(v) = 3$, $v$ has at most one neighbour in $C$, a contradiction to $C$ being a $K_3$.
\end{proof}
\paragraph{}
We proceed by a discharging argument. For each vertex $v \in V(G)$ set the initial charge as
$$ch_0(v) = d(v) - 4$$
and for each edge $e \in E(G)$ set the initial charge
$$ch_0(e) = 1/3.$$
Then our total charge is, using handshaking,
\begin{align*}
\sum_{v\in V(G)} ch_0(v) + \sum_{e \in E(G)} ch_0(e) &= \sum_{v \in V(G)} d(v) - 4|V(G)| + \frac{1}{3} |E(G)|\\ &=  \frac{7}{3}|E(G)| - 4|V(G)\\ &< \frac{35}{9}|V(G)| - \frac{36}{9}|V(G)|\\ &< 0
\end{align*}
with the first inequality following from our assumption that $|E(G)| < 5|V(G)|/3$.
\paragraph{}
Redistribute the charge, to obtain $ch_1$, according to the following Rules:
\begin{enumerate}
	\item If edge $uv$ contains precisely one vertex $u$ such that $d(u) = 3$ then $uv$ sends $1/3$ charge to $u$.
	\item If edge $uv$ is such that $d(u) \geq 4$ and $d(v) \geq 4$ and $u,v$ are in a triangle with vertex $w$ then $uv$ sends $1/3$ charge to $w$.
\end{enumerate}
\paragraph{}
Since vertices sends no charge, clearly every vertex $v \in V(G)$ with $d(v) \geq 4$ has $$ch_1(v) = ch_0(v) \geq 0.$$ Now we will show that if $v$ is a vertex with $d(v) = 3$ then $ch_1(v) \geq 0$. From Claim \ref{1.1} $v$ is adjacent with at most one vertex of degree $3$. If $v$ is adjacent with no degree three vertices then each edge incident with $v$ sends $\frac{1}{3}$ charge to $v$ by Rule $1$ and hence
$$ch_1(v) = ch_0(v) + 3\cdot\frac{1}{3} = 0.$$
On the other hand, if $v$ is adjacent with one degree three vertex then $v$ receives $\frac{2}{3}$ charge from incident edges by Rule $1$. The remaining needed $\frac{1}{3}$ is obtained from Rule $2$ as follows. Since $v$ is adjacent with a degree three vertices, $v$ lies in a triangle, and since every triangle contains at most one degree three vertex the edge containing the other two vertices of this triangle sends $\frac{1}{3}$ charge to $v$. Therefore $ch_1(v) \geq 0$.
\paragraph{}
It remains to verify that each edge $uv$ satisfies $ch_1(uv) \geq 0$. Edge $uv$ can satisfy at most one of Rule $1$ and Rule $2$. If $uv$ satisfies Rule $1$ it sends $\frac{1}{3}$ charge and so $ch_1(uv) = 0$. Instead if $uv$ satisfies Rule $2$ it can only do so for one triangle, since no two distinct triangles share a common edge. So $uv$ only sends $\frac{1}{3}$ charge by Rule $2$. Therefore $ch_1(uv) \geq 0$.
\paragraph{}
Therefore we have a contradiction as our total charge is negative but $ch_1$ of each object is non-negative.$\blacksquare$
\end{document}