### 2018-02-17 CO749-GraphColoring/assignments/A3.tex
2ab

\documentclass[letterpaper,12pt,oneside,onecolumn]{article}
\usepackage[margin=1in, bottom=1in, top=1in]{geometry} %1 inch margins
\usepackage{amsmath, amssymb, amstext}
\usepackage{fancyhdr}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{theorem}
\usepackage{tikz}
\usepackage{tkz-berge}
\usepackage[braket, qm]{qcircuit}
\usepackage{hyperref}

%Macros
\newcommand{\A}{\mathbb{A}} \newcommand{\C}{\mathbb{C}}
\newcommand{\D}{\mathbb{D}} \newcommand{\F}{\mathbb{F}}
\newcommand{\N}{\mathbb{N}} \newcommand{\R}{\mathbb{R}}
\newcommand{\T}{\mathbb{T}} \newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}

\newcommand{\cA}{\mathcal{A}} \newcommand{\cB}{\mathcal{B}}
\newcommand{\cC}{\mathcal{C}} \newcommand{\cD}{\mathcal{D}}
\newcommand{\cE}{\mathcal{E}} \newcommand{\cF}{\mathcal{F}}
\newcommand{\cG}{\mathcal{G}} \newcommand{\cH}{\mathcal{H}}
\newcommand{\cI}{\mathcal{I}} \newcommand{\cJ}{\mathcal{J}}
\newcommand{\cK}{\mathcal{K}} \newcommand{\cL}{\mathcal{L}}
\newcommand{\cM}{\mathcal{M}} \newcommand{\cN}{\mathcal{N}}
\newcommand{\cO}{\mathcal{O}} \newcommand{\cP}{\mathcal{P}}
\newcommand{\cQ}{\mathcal{Q}} \newcommand{\cR}{\mathcal{R}}
\newcommand{\cS}{\mathcal{S}} \newcommand{\cT}{\mathcal{T}}
\newcommand{\cU}{\mathcal{U}} \newcommand{\cV}{\mathcal{V}}
\newcommand{\cW}{\mathcal{W}} \newcommand{\cX}{\mathcal{X}}
\newcommand{\cY}{\mathcal{Y}} \newcommand{\cZ}{\mathcal{Z}}

\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}

\newenvironment{proof}{{\bf Proof:  }}{\hfill\rule{2mm}{2mm}}
\newenvironment{proofof}[1]{{\bf Proof of #1:  }}{\hfill\rule{2mm}{2mm}}
\newenvironment{proofofnobox}[1]{{\bf#1:  }}{}\newenvironment{example}{{\bf Example:  }}{\hfill\rule{2mm}{2mm}}

%\renewcommand{\thesection}{\lecnum.\arabic{section}}
%\renewcommand{\theequation}{\thesection.\arabic{equation}}
%\renewcommand{\thefigure}{\thesection.\arabic{figure}}

\newtheorem{fact}{Fact}[section]
\newtheorem{lemma}[fact]{Lemma}
\newtheorem{theorem}[fact]{Theorem}
\newtheorem{definition}[fact]{Definition}
\newtheorem{corollary}[fact]{Corollary}
\newtheorem{proposition}[fact]{Proposition}
\newtheorem{claim}[fact]{Claim}
\newtheorem{exercise}[fact]{Exercise}
\newtheorem{note}[fact]{Note}
\newtheorem{conjecture}[fact]{Conjecture}

\newcommand{\size}[1]{\ensuremath{\left|#1\right|}}
\newcommand{\ceil}[1]{\ensuremath{\left\lceil#1\right\rceil}}
\newcommand{\floor}[1]{\ensuremath{\left\lfloor#1\right\rfloor}}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
\DeclarePairedDelimiter\norm{\lVert}{\rVert}%
%END MACROS
%Page style
\pagestyle{fancy}

\listfiles

\raggedbottom

\lhead{\today}
\rhead{W. Justin Toth - A3} %CHANGE n to ASSIGNMENT NUMBER ijk TO COURSE CODE
\renewcommand{\headrulewidth}{1pt} %heading underlined
%\renewcommand{\baselinestretch}{1.2} % 1.2 line spacing for legibility (optional)

\begin{document}
\section{}
A $k$-critical graph $G$ is $m$-\emph{splittable} (with $2\le m \le k-1$) if for every vertex $z\in V(G)$ and every split of $z$ into vertices $z_1,\ldots, z_m$ of positive degree , the resulting graph has a $(k-1)$-coloring such that $z_1, \ldots, z_m$ have distinct colors. We say such $G$ is \emph{splittable} if $G$ is $m$-splittable for every $m$ such that $2\le m \le k-1$.
\begin{lemma}
	An Ore-composition of two splittable $k$-critical graphs is splittable $k$-critical.
\end{lemma}
\begin{proof}
	Let $G_1$ and $G_2$ be two splittable $k$-critical graphs. Let $G$ be the Ore-composition of $G_1$ and $G_2$ obtained by deleting edge $xy \in E(G_1)$ and splitting vertex $z \in V(G_2)$ into $z_1$ and $z_2$ of positive degree and identifying $x$ with $z_1$ and $y$ with $z_2$. We denote the identified vertex for $x$ with $z_1$ as $z_x$, and for $y$ with $z_2$ as $z_y$ in $G$.
	\paragraph{}
	We first show that $G$ is $k$-critical. Suppose for a contradiction $G$ has a $(k-1)$-coloring $\phi$. If $\phi(z_x) \neq \phi(z_y)$ then $\phi$ restricted to $G_1$ (taking in the obvious manner $\phi(x) = \phi(z_x)$ and $\phi(y) = \phi(z_y)$) is a $(k-1)$-coloring of $G_1$. If $\phi(z_x) = \phi(z_y)$ then $\phi$ restricted to $G_2$, and takiing $\phi(z) = \phi(z_x) = \phi(z_y)$ yields a $(k-1)$-coloring of $G_2$. In either case we contradicted that $G_1$ and $G_2$ do not have a $(k-1)$-coloring. Therefore $G$ is not $(k-1)$-colorable.
	\paragraph{}
	To finish verifying that $G$ is $k$-critical we need to show that every strict subgraph of $G$ is $(k-1)$-colorable. Let $H \subseteq G$ be a strict subgraph of $G$. We proceed by case-distinction on $H$.
		\subparagraph{} Case $1$: $z_x$ is not contained in $H$ and $z_y$ is not contained in $H$. Then $H$ is the disjoint union of a strict subgraph of $G_1$ and a strict subgraph of $G_2$. Since $G_1$ and $G_2$ are $k$-critical we can obtain a $(k-1)$-coloring of their respective strict subgraphs in $H$ and combine them to obtain a $(k-1)$-coloring of $H$.
		\subparagraph{} Case $2:$ Exactly one of $z_x, z_y$ is in $H$. Without loss of generality say $z_x \in V(H)$ and $z_y \not\in V(H)$. Then treating $z_x$ as $x$ in $G_1$ and $z$ in $G_2$ we see that $H$ is the union of two strict subgraphs $H_1 \subseteq G_1$ and $H_2 \subseteq G_2$, overlapping only at $z_x$. We see $H_2$ is a strict subgraph since the degree of $z_1$ and $z_2$ are both positive. Again we invoke criticality of $G_1$ and $G_2$ to obtain $(k-1)$-colorings of $H_1$ and $H_2$, which we join into a $(k-1)$-coloring of $H$. Note we permute the coloring of $H_1$ so the two colorings agree at $z_x$ if necessary.
		\subparagraph{} Case $3:$. Both $z_x$ and $z_y$ are in $H$. Let $G_2'$ denote the graph $G_2$ with $z$ split into $z_1$ and $z_2$. Then $H$ is the union of two subgraphs $H_1 \subseteq G_1$ and $H_2 \subseteq G_2'$ overlapping only at $\{z_x,z_y\}$. Since $H$ is a strict subgraph of $G$ one of $H_1$ or $H_2$ is a strict subgraph of $G_1-xy$ or $G_2'$ respectively.
		\subparagraph{} First suppose $H_1$ is a strict subgraph of $G_1-xy$. Then invoking the criticality of $G_1$, we obtain that $H_1 + xy$ has a $(k-1)$-coloring $\phi_1$. Then $\phi_1$ is a $(k-1)$-coloring of $H_1$ where $\phi_1(x)\neq \phi_1(y)$. Since $G_2$ is $2$-splittable, there exists a $(k-1)$-coloring $\phi_2$ of $G_2'$. Note that $\phi_2(z_1) \neq \phi_2(z_2)$ for otherwise setting $\phi_2(z) = \phi_2(z_1) = \phi_2(z_2)$ yields a $(k-1)$-coloring of $G_2$. Hence we may permute the colors of $\phi_2$ so $\phi_2(z_1) = \phi_1(x)$ and $\phi_2(z_2) = \phi_1(y)$. Then the coloring obtained by combining $\phi_1$ and $\phi_2$ is a $(k-1)$-coloring of $H$.
		\subparagraph{}
		So we may assume $H_1$ is not a strict subgraph of $G_1-xy$, i.e. $H_1 = G_1-xy$. Then $H_2$ is a strict subgraph of $G_2'$. By identifying $z_x$ and $z_y$ as $z$ in $H_2$ we obtain a strict subgraph of $G_2$. By criticality of $G_2$, this subgraph has a $(k-1)$-coloring $\phi_2$. By setting $\phi_2(z_x) = \phi_2(z_y) = \phi_2(z)$ we obtain a $(k-1)$-coloring of $H_2$ wherein $z_x$ and $z_y$ receive the same color. Invoking criticality of $G_1$, we obtain a $(k-1)$-coloring $\phi_1$ of $H_1$ ($=G_1-xy$). Observe that $\phi_1(x) = \phi_1(y)$, for otherwise $\phi_1$ is a $(k-1)$-coloring of $G_1$. But then we may permute the colors of $\phi_2$ so $\phi_2(z_1) = \phi_1(x)$ and $\phi_2(z_2) = \phi_1(y)$. Again the coloring obtained by combining $\phi_1$ and $\phi_2$ is a $(k-1)$-coloring of $H$, as desired.
	\paragraph{Splittability}
	Thus we have shown that $G$ is $k$-critical. It remains to verify that $G$ is splittable. First consider $v \in V(G) \backslash\{z_x,z_y\}$. Consider $G'$, where we split $v$ into $v_1, \dots, v_m$ each with positive degree, $2 \leq m \leq k-1$. If $v \in V(G_1)$ then apply the corresponding split $v_1, \dots, v_m$ to $G_1$, and use that $G_1$ is splittable to obtain a $(k-1)$-coloring $\phi_1$. Observe $\phi_1(x) \neq \phi_1(y)$ since $xy \in E(G_1)$. Since $G_2$ is $2$-splittable, $G - G_1$ has a $(k-1)$-coloring $\phi_2$. Since $G_2$ is not $(k-1)$-colorable, $\phi_2(z_x) \neq \phi_2(z_y)$. Again we may permute colors of $\phi_2$ so $\phi_1$ and $\phi_2$ agree on $z_x$ and $z_y$, obtaining a $(k-1)$-coloring of $G'$.
	\paragraph{}
	If $v \in V(G_2)$ then apply the corresponding split $v_1, \dots, v_m$ to $G_2$ and use splittability to obtain a $(k-1)$-coloring $\phi_2$. Setting $\phi_2(z_x) = \phi_2(z_y) = \phi_2(z)$ we obtain a $(k-1)$-coloring of $G-G_1$. By criticality, $G_1-xy$ has a $(k-1)$-coloring $\phi_1$. Since $G_1$ is not $(k-1)$-colorable, $\phi_1(x) = \phi_1(y)$. Hence we may permute colors of $\phi_2$ so $\phi_1$ and $\phi_2$ agree on $z_x$ and $z_y$, obtaining a $(k-1)$-coloring of $G'$.
	\paragraph{}
	It remains to consider splitting one $z_x$ or $z_y$ in $G$. Without loss of generality let $G'$ be the graph obtained by splitting $z_x$ into $z_x^1, \dots, z_x^m$, each with positive degree, $2 \leq m \leq k-1$. Split $x$ in $G_1$ into $x^1, \dots, x^m$ with neighbourhoods $N(x^i) = N(z_x^i)\cap V(G_1)$ yielding graph $G_1'$. Obtain a $(k-1)$-coloring $\phi_1$ of $G_1'$ by splittability of $G$.
	\subparagraph{}Case $1:$ $\phi_1(x^i) = \phi_1(y)$ for some $i \in \{1,\dots, m\}$. Split $z$ in $G_2$ into $z^1,\dots, z^m$ with neighbourhoods $N(z^j) = N(z_x^j) \cap V(G_2)$ for $j \neq i$ and $$N(z^i) = (N(z_x^i) \cup N(z_y)) \cap V(G_2)$$
	obtaining graph $G_2'$. Color $G_2'$ by splittability of $G_2$ obtaining $(k-1)$-coloring $\phi_2$. Setting $\phi_2(z_y) = \phi_2(z^i)$ we obtain a $(k-1)$-coloring of $(G-G_1) \cup \{z_x^1, \dots, z_x^m, z_y\}$. Since $\phi_1$ and $\phi_2$ give all the split vertices different colors, and give $z_y$ the same color $z_x^i$ we may permute $\phi_1$ and $\phi_2$ to agree on $\{z_x^1, \dots, z_x^m, z_y\}$ obtaining a $(k-1)$-coloring of $G'$.
	\subparagraph{}Case $2:$ $\phi_1(x^i) \neq \phi_2(y)$ for all $i \in \{1,\dots, m\}$. Notice in this case $m < k-1$ by the pidgeonhole principle since $\phi_1$ assigns different colors to each split vertex. Split $z$ in $G_2$ into $z^1, \dots, z^m, z^{m+1}$ with neighbours $N(z^i) = N(z_x^i) \cap V(G_2)$ for $i \leq m$ and set $N(z^{m+1}) = N(z_y) \cap V(G_2)$. Applying splitability we obtain a $(k-1)$-coloring $\phi_2$. Now both $\phi_1$ and $\phi_2$ give $\{z_x^1, \dots, z_x^m, z_y\}$ different colors, so we may permute their colors to agree on $\{z_x^1, \dots, z_x^m, z_y\}$, and join the colorings to obtain a coloring of $G'$, as desired.
	\end{proof}
\begin{corollary}
	If $G$ is a $k$-Ore graph then $G$ is splittable $k$-critical.
\end{corollary}
\begin{proof}
	Suppose not. Let $G$ be a minimum counterexample. Then $G$ cannot be the Ore-composition of two $k$-Ore graphs $G_1$, $G_2$, for otherwise $G_1$ and $G_2$ are splittable $k$-critical by minimality and hence $G$ is splittable $k$-critical by the previous Lemma.
	\paragraph{}
	Thus $G$ is isomorphic to $K_k$. It remains to show $K_k$ is splittable. Suppose we want to split vertex $v \in V(G)$ into $v_1, \dots, v_m$ with $2\leq m \leq k-1$ such that each $v_i$ has positive degree, obtaining graph $G'$. We can trivially color $G - v$ with $k-1$ colors, obtaining coloring $\phi$. Now we can extend $\phi$ to $G'$ since each $v_i$ has degree at most $k-2$ and hence $|\phi(N(v_i))| \leq k-2$ and there is a free color available to assign $v_i$. This contradicts that
\end{proof}

\section{}
\subsection{a}
\begin{lemma}
	If $G$ is the Ore-composition of two graphs $G_1$ and $G_2$ then $\alpha(G) \leq \alpha(G_1) + \alpha(G_2)$.
\end{lemma}
\begin{proof}
Let $G$ be the Ore-composition of graphs $G_1$ and $G_2$ obtained by deleting edge $xy \in E(G_1)$ and splitting vertex $z \in V(G_2)$ into $z_1$ and $z_2$ of positive degree and identifying $x$ with $z_1$ and $y$ with $z_2$. We denote the identified vertex for $x$ with $z_1$ as $z_x$, and for $y$ with $z_2$ as $z_y$ in $G$. Let $S \subseteq V(G)$ be an independent set in $G$ such that $|S| = \alpha (G)$. 
Then treating $z_x$ as $x$ and $z_y$ as $z$ we can partition $S$ into two independent sets $S_1\subseteq V(G_1)$ and $S_2\subseteq V(G_2)$ as $S_1 = V(G_1) \cap S$ and $S_2 = S\backslash S_1$. Thus
$$\alpha(G) = |S| = |S_1| + |S_2| \leq \alpha(G_1) + \alpha(G_2)$$
as desired.
\end{proof}

\subsection{b}
\begin{lemma}
	If $G$ is $k$-Ore then $\alpha(G) = \frac{v(G) - 1}{k-1}$.
\end{lemma}
\begin{proof}
	We proceed by induction on $v(G)$. The smallest $k$-Ore graph is $K_k$, for which
	$$\alpha(K_k) = 1 = \frac{k-1}{k-1} = \frac{v(G) -1}{k-1}.$$
	So the base case for induction is satisfied.
	\paragraph{}
	So we consider a $k$-Ore graph $G$ that is the Ore-composition of $G_1$ and $G_2$ and suppose for induction that
	$$\alpha(G_1) = \frac{v(G_1) - 1}{k-1} \quad\text{ and }\quad \alpha(G_2) = \frac{v(G_2) -1}{k-1}.$$
	We observe from the previous Lemma, and induction, that
	$$\alpha(G) \leq \alpha(G_1) + \alpha(G_2) = \frac{v(G_1) +v(G_2)- 2}{k-1} = \frac{v(G) + 1 - 2}{k-1} = \frac{v(G) - 1}{k-1}.$$
	Thus $\alpha(G) \leq \frac{v(G) - 1}{k-1}$. To show $\alpha(G) \geq \frac{v(G) - 1}{k-1}$ it suffices to construct an independent set in $G$ of size $\frac{v(G) - 1}{k-1}$.
	\paragraph{}
	For $i \in \{1, \dots, 2\}$ let $S_i$ be in an independent set in $G_i$ satisfying $$|S_i| = \alpha(G_i) = \frac{v(G_i) - 1}{k-1}.$$ We use the same notion as the above problems for $x,y, z, z_x, z_y$. Since $xy \in E(G_1)$, $|\{x,y\} \cap S_1| \leq 1$. Without loss of generality suppose $y \not\in S_1$. Treating $x$ as $z_x$ and $z$ as $z_y$ we observe that
	$$S := S_1 \cup S_2$$
	is an independent set of $G$ of size $|S_1| + |S_2|$. But then we see that
	$$|S| = |S_1| + |S_2| =  \frac{v(G_1) +v(G_2)- 2}{k-1} = \frac{v(G) - 1}{k-1}.$$
	Thus $\alpha(G) \geq \frac{v(G) - 1}{k-1}$, and therefore by induction the Lemma is proven.
\end{proof}
\end{document}