### 2019-03-08 CUT/paper.tex
Research Paper outline

\documentclass{article}

\usepackage{apacite}

\begin{document}
\begin{abstract}
Research Question: ``How does assessment format impact student learning outcomes, and instructor measurement thereof''.

New Research Question: ``Can innovation assessment formats improve student learning outcomes, and an instructors capacity to measure their achievement.''
\end{abstract}

\section{Introduction}

\section{Particulars of Written Finals}
\subsection{Question Formats}
\subsection{Time Pressure}
\subsection{Open Book}
%can include takehome

\section{Oral Exams}
\subsection{Implementation}
\subsection{Results}
\subsection{Benefits}
\subsection{Drawbacks}
\subsection{Data}

\section{Collaborative Exams}
\subsection{Group}
\subsection{Two-Stage}

\section{Student Showcase}
\subsection{Portfolios}
\subsection{Student-Written Exam Problems}

\section{Stakeholder Impressions}
%how do students and instructors perceive different exam formats?

\section{Recommendations}
\subsection{Designing Exams to Encourage Good Learning Habits}
\subsection{Assessing a Variety of Cognitive Processes}
\subsection{Encouraging Genuine Engagement}
\nocite{*}
\bibliographystyle{apacite}
\bibliography{references.bib}

\end{document}